### 2020-08-23 paper/appendix.Rmd
init

---
title: "Appendix M"
output:
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
    fig_width: 4
    fig_height: 4
    fig_caption: true
    highlight: tango
---
<!-- Run this to compile:
R -e "rmarkdown::render('appendix.Rmd', rmarkdown::pdf_document(), output_file='appendix.pdf', clean=TRUE)"
-->

# Basic / Classical RMT with Empirical Data

## Requirements and Overview
All you *really* need for basic RMT analysis is either:

1. a large (1000x1000) symmetric square matrix
2. a large sample of smaller symmetric square matrices
   - e.g. 10 or more 100x100 symmetric matrices

Provided your data / system has $N$ entities, and each entity has a
series of scalar values associated with them (e.g. you have $S_1, S_2, \dots, S_N$
where $S_i = \{t_{i_1}, ...\;, t_{i_{M_i}}\}$, $t_{i_j} \in \mathbb{R}$),
you can always generate a symmetric matrix with any function
that symmetrically compares the two series. I.e. if we have
$F$ such that $F(x, y) = F(y, x)$ for real series $x$ and $y$,
then the matrix $[F(S_i, S_j)]$ will be symmetric. Sensible choices of
$F$ might be Pearson's correlation coefficient, mutual information,
cross-correlation (convolution), or other distance / similarity metrics.

However, if you want to be able to actually find anything
interpretable (or interesting), you probably want a system that can be
roughly modeled as a number of somewhat distinct but interacting (or
similar, or connected) entities. E.g. particles bouncing off each
other, energy levels of a Uranium atom, distances between cars on a
freeway or in a parking lot, stock prices in a stock market, etc. It
is not clear how one would make sense of an RMT analysis of the
correlations of a static system (e.g. correlations between voxels or
ROIs of a structural MRI image).

The standard RMT procedure is then to compute some curves (metrics),
and compare (usually by eye) those curves to those predicted by RMT.
Typically, empirical results are compared to only the Gaussian
Orthogonal Ensemble (GOE) case, and/or the Poisson case, since each of
these represent two interpretational extremes.

The GOE is essentially a matrix where each entry is sampled from
the standard normal distribution. More precisely, it is the set of random
matrices:

$$A = [a_{ij}] \quad\mathrm{where}\quad A = A^T, \quad a_{ij} ~
\begin{cases}
  N(0, 1) &\mathrm{if}\; i \ne j\\
  \sqrt{2}N(0, 1) &\mathrm{if \; i = j}\\
\end{cases}
$$

If your matrix or matrices look like they are coming from a GOE
ensemble, you conclude the interactions in your system are highly
chaotic/random/noisy, i.e. that there is no real information (in the
mathematical and statistical sense) to be extracted there. This is a
pretty big deal, so if you see GOE behaviour, that's important.

The Poisson case is less clear to me currently, but it represents a
different kind of physical / mathematical extreme, and seems to represent
a more *regular* system. E.g. harmonic oscillators (where a particle's
position depends only on the field and its current position, and
not the positions of any other particles) produce some RMT metrics
similar to Poisson systems (Guhr & Mehta, 1998).

## Computational Approach
1. Compute a symmetric correlation, covariance, similarity, distance, mutual
   information, convolution, etc. matrix for your system
2. Compute the eigenvalues of that matrix
3. Sort the eigenvalues in ascending order
4. Compute the eigenvalue step / counting function
5. Unfold the eigenvalues
   - i.e. fit a smooth curve from the eigenvalues to the step function
   - use the predicted eigenvalues from this fit as the new unfolded values
6. Compute the spectral observables (curves) of the unfolded eigenvalues
7. Compare the computed curves to those predicted by theory

# Equations
## Fundamentals
### Dirac Delta Function

The function $\delta(x-t)$, *Dirac's delta* is informally defined, for constant $t$ as:

$$\delta(x-t) =
\begin{cases}
  \infty & \quad \text{if  } x = t \\
  0 & \quad \text{if  } x \ne t
\end{cases}$$

and where $\int_{-\infty}^{\infty}\delta(x - t) \;\mathrm{d}x = 1$.
Mostly it is a clever tool to place an infinitely thin but smooth
"spike" at a value to allow discrete-valued functions or sums to be
written continuously. It can be justified formally (as an entity used
only in integrals, and as the limit of a series of increasingly narrowing
bump functions with fixed area under the curve).

## Eigenvalue Step Function
Suppose we have our ordered eigenvalues $\{\lambda_1, ...,
\lambda_N\}$. Then $\eta(\lambda)$, the number of eigenvalues on
$[-\infty, \;\lambda]$, the eigenvalue **step** or **staircase
function** is:

$$\eta(\lambda) = \int_{-\infty}^{\lambda}
\sum_{n=1}^{N}\delta(\lambda - \lambda_n) \;\mathrm{d}\lambda \;\; =
\;\; \text{the numbers eigenvalues } \lambda_i \text{ less than
}\lambda$$

![Step Function for N=10000 GOE Matrix](figures/step_function_zoomed_out.png)
**Figure 1.** Example step function from a N=10 000 GOE matrix:

![Step Function for N=10000 GOE Matrix](figures/step_function_zoomed_in.png)
**Figure 2.** Step function fine structure.

In code:
```python
def stepFunction(eigs: np.array, x: np.array) -> np.array:
    if x == eigs:
        return np.arange(0, len(eigs) + 1)  # shortcut, return [0, 1, ... len(eigs)]
    steps = np.empty(len(eigs))
    for i, val in x:
        steps[i] = len(eigs[eigs <= val])  # count eigs <= each x
    return steps
```

## Unfolding
In order to compare the eigenvalues observed from a system to those
predicted from theory, we have to "unfold" the eigenvalues so that
they exist on the same scale / as those of the classical RMT ensembles.
This is a crucial and necessary step, and can be thought of as roughly
analagous to denoising (paraphrasing Guhr & Mehta, it is "removing the
system-specific mean levels to allow comparison of the unversal
components").

To unfold, we decompose the step function--in practice, by fitting a
polynomial or spline to $\{\lambda, \eta(\lambda)\}$--into a smooth
part $\tilde\eta(\lambda)$ and fluctuating part
$\eta_{\mathrm{fl}}(\lambda)$ such that

$$\eta(\lambda) = \tilde\eta(\lambda) + \eta_{\mathrm{fl}}(\lambda)$$

In certain special cases / systems where the physical model is fully
specified and understood, the functional forms of these smooth and
fluctuating parts can be calculated precisely. But for novel systems
being investigated with RMT, especially when there is no clear or
explicit underlying model, this unfolding procedure is essentially
guesswork.

In practice, it is most common to fit a smooth spline, or polynomial
of degree 5-11 $f$ to $\{\lambda, \eta(\lambda)\}$, and hope that $f$
is a good approximation to $\tilde\eta$. We then define

$$\{u_i = f(\lambda_i) \;\; | \;\; i = 1, 2, ..., N\}$$

to be the unfolded eigenvalues.

### Unfolding Algorithm
The essence of the unfolding process is a simple fitting procedure.

```python
def polynomial_unfold(eigs, degree):
    steps = np.empty([len(eigs)])  # allocate array for step function values
    for i, eig in enumerate(eigs):
        steps[i] = len(eigs[eigs < eig])  # count eigenvalues less than eig
    # steps = np.arange(0, len(eigs) + 1)  # alternate faster computation
    fit = polyfit(eigs, steps, degree)
    unfolded = polyval(eigs, fit)
    return unfolded
```
However, in practice the large eigenvalues of the system may act as anchor
points during fitting, heavily destroying the ability of the unfolding
process to extract information from the system. These large eigenvalues
must be dealt with in some manner (trimmed, fit separately). Additionally,
more complex unfolding procedures are also possible (using detrending, signal
processing approaches, and more local fitting).

### Unfolding Issues
* Long-range spectral observables are wildly sensitive to the unfolding procedure.
* The unfolding procedure is extremely arbitrary
  * choice of fitting procedure
    * e.g. polynomial, spline, EMD, other signal processing methods, detrending or not
  * fitting parameters
    * e.g. degree of polynomial / number of knots or type of spline, other fitting parameters
    * proportion of eigenvalues to fit (e.g. trimming eigenvalues of "large" absolute value)
    * fit accuracy (e.g. in theory we fit a continuous function $\eta(\lambda)$, but in practice
      we only evaluate $\eta(\lambda)$ on some grid of our choice
  * no way to test validity for unknown ensembles / systems
    * i.e. bad unfolding can make GOE systems look not GOE, or non-GOE systems look GOE (!!)

# Spectral Observables (Empirical Tests of RMT)
The spectral observables are the statistics of RMT. They are
statistics in the formal / general sense, i.e. they summarize random
variables. However, there are very, very few statistical tests in RMT.
One compares empirical results with RMT theory in the usual way
physicists do, e.g. via "binocular inspection".

## Nearest-Neighbour Spacing Distribution (Short Range)
Given observed eigenvalues $\{\lambda_1 \le ... \le \lambda_N\}$, we define

$$S_i = \lambda_{i+1} - \lambda_{i}, \qquad i = 1, 2, \dots N-1$$

the distances between the eigenvalues, and

$$D = \langle S_i \rangle$$

the mean distance between them. Then the (relative) level spacings are

$$s_i = S_i / D.$$

From theory, we know $p(s)$, is the probability of the spacings
between nearest neighboring eigenvalues. For certain cases (matrix
ensembles, or uncorrelated / Poisson case) the specific form of the
distribution is known. By computing the histogram (or kernel density
estimate) based on the empirically observed $s_i$ calculated from the
unfolded eigenvalues, we can compare our observations to the distribution
expected from theory.

For the Poisson case:

$$p(s) = \mathrm{exp}(-s)$$

For the GOE, $\beta = 1$,

$$p_1(s) = \frac{\pi}{2}s\;\mathrm{exp}\left(-\frac{\pi}{4}s^2\right)$$

For the GUE, $\beta = 2$,

$$p_2(s) = \frac{32}{\pi^2}s\;\mathrm{exp}\left(-\frac{4}{\pi}s^2\right)$$

For the GSE, $\beta = 4$,

$$p_4(s) = \frac{262144}{729\pi^3}s\;\mathrm{exp}\left(-\frac{64}{9\pi}s^2\right)$$

Note the domain of $p(s)$ is always $[0, \infty]$.

![NNSD](figures/nnsd_curves.png)
**Figure.** Predicted and actual distributions from a 1000x1000 GOE matrix.

## Level Number Variance $\Sigma^2(L)$

Let $N(L, c)$ count the number of eigenvalues on the interval $[c,\;c+L]$ "on the unfolded scale", i.e. $u_1 = \tilde\eta(\lambda_1)$. I.e. if we had the eigenvalues $E = \{1, 2, 3, 4.5, 4.7, 5, 6, 7, 8\}$, then:

$$
\begin{aligned}
  N(3, 3) \;&=\; |\{\lambda \in E : \lambda \in [3, 6]| &=& \quad|\{3, 4.5, 4.7, 5, 6\}| &= 5 \\
  N(3, 1.6) \;&=\; |\{\lambda \in E : \lambda \in [3, 4.6]| &=& \quad|\{3, 4.5\}| &= 2 \\
  N(100, 7) \;&=\; |\{\lambda \in E : \lambda \in [7, 107]| &=& \quad|\{7, 8\}| &= 2 \\
  N(100, 7) \;&=\; |\{\lambda \in E : \lambda \in [7, 107]| &=& \quad|\{7, 8\}| &= 2 \\
\end{aligned}
$$

Then we define:

$$\Sigma^2(L) = \left\langle N(L, c)^2 \right\rangle_c -  \left\langle N(L,c) \right\rangle_c^2,$$

the level number variance.

### $\Sigma^2(L)$ Algorithm
Algorithmically, for a particular value of $L$, we might compute $\Sigma^2$ like:
```python
def sigma_iter(unfolded: np.array, L: float, c_iters: int = 100):
    # allocate arrays to save counts of levels and square of those counts
    levels = np.empty((c_iters), dtype=np.float64)
    levels_sq = np.empty((c_iters), dtype=np.float64)

    # randomly choose c within the unfolded scale, and repeat c_iters times
    for i in range(c_iters):
        c_start = np.random.uniform(np.min(unfolded), np.max(unfolded))
        start, end = c_start - L/2, c_start + L/2
        # count number of eigenvalues in [start, end]
        n_within = len(unfolded[(unfolded >= start) & (unfolded <= end)])
        levels[i], levels_sq[i] = n_within, n_within**2

    ave = np.mean(levels)
    av_of_levels_sq = ave * ave
    av_of_sq_levels = np.mean(levels_sq)
    return av_of_sq_levels - av_of_levels_sq
```

For the classical ensembles and large $L$,
$\Sigma^2$, $\gamma = 0.5772\dots$ Euler's constant, $\Sigma^2$ has
approximations (valid to order $1/L$):

$$
\begin{aligned}
 \Sigma^2_\mathrm{Poisson}(L) =&\;L \\
 \Sigma^2_\mathrm{GOE}(L) =&\; \frac{2}{\pi^2}\left( \mathrm{ln}(2\pi L) + \gamma + 1 - \frac{\pi^2}{8} \right) \\
 \Sigma^2_\mathrm{GUE}(L) =&\; \frac{1}{\pi^2}\left( \mathrm{ln}(2\pi L) + \gamma + 1 \right) \\
 \Sigma^2_\mathrm{GSE}(L) =&\; \frac{1}{2\pi^2}\left( \mathrm{ln}(4\pi L) + \gamma + 1 + \frac{\pi^2}{8} \right) \\
\end{aligned}
$$

![Level Variance](figures/levelvar_curves.png)
**Figure.** Predicted and actual curves from a 1000x1000 GOE matrix.

There are also precise forms that are more accurate for small $L$.

## Spectral Rigidity $\Delta_3(L)$
If $\langle f \rangle_{t}$ denotes the average of $f$ over all $t$, and
$\hat{\eta}(\xi)$ represents the the number of levels on the unfolded scale, then

\begin{equation}
  \label{eq:delta3}
  \Delta_3(L) = {\left\langle \min_{A, B} \frac{1}{L} \int_{c}^{c+L}(\hat{\eta}(\lambda) -A\lambda - B)^2 \;\mathrm{d}\lambda\right\rangle}_{c}
\end{equation}

![Spectral Rigidity](figures/rigidity_calculation.png)
**Figure.** Average deviation of a section of the staircase function from a straight line.

is the spectral rigidity. Note that $A$ and $B$ may be different for each choice of $c$, $L$. For the motivation, note that if $(\hat{\eta}(\lambda) -A\lambda - B)^2 = 0$ then $\hat\eta$ is linear, i.e. the staircase function is extremely rigid. Also note that if $y = A\lambda + B$ is the least squares linear fit to $\hat\eta(\lambda)$ on $[c, \;c+L]$, then $A$ and $B$ minimize the integral.

### $\Delta_3(L)$ Algorithm
Algorithmically then, for a particular $L$ and $c$ value, we might calculate $\Delta_3(L)$ as:

```python
def spectralIter(eigs: np.array, unfolded: np.array, L, c_start, interval_gridsize) -> float:
    # c_start is in space of eigs
    grid = np.linspace(c_start - L/2, c_start + L/2, interval_gridsize)
    # compute step function values for each value in grid
    steps = stepFunction(eigs, grid)
    A, B = computeLinearFit(grid, steps)  # get slope A and intercept B from fit
    fit = A*grid + B                      # fit the line to the step function
    y = (steps - fit)**2                  # compute squared deviation of fit
    delta3 = np.trapz(y, grid)            # numerically integrate the values
    return delta3 / L
```
choosing $c$ values randomly and as many times as desired for stability / precision, for each $L$.

For the classical ensembles, the shape of $\Delta_3$ is known, and can be approximated to order $1/L$ by:

$$
\begin{aligned}
 \Delta_{3,\mathrm{Pois}son}(L) =&\;L/15 \\
 \Delta_{3,\mathrm{GOE}}(L) =&\; \frac{1}{\pi^2}\left( \mathrm{ln}(2\pi L) + \gamma - \frac{5}{4} - \frac{\pi^2}{8} \right) \\
 \Delta_{3,\mathrm{GUE}}(L) =&\; \frac{1}{2\pi^2}\left( \mathrm{ln}(2\pi L) + \gamma - \frac{5}{4} \right) \\
 \Delta_{3,\mathrm{GSE}}(L) =&\; \frac{1}{4\pi^2}\left( \mathrm{ln}(4\pi L) + \gamma - \frac{5}{4} + \frac{\pi^2}{8} \right) \\
\end{aligned}
$$

![Spectral Rigidity](figures/rigidity_curves.png)
**Figure.** Predicted Spectral Rigidity curves, with points from actual GOE data.

# Theory
There are two ways to look at RMT and justify its applications in
fMRI: mathematical and physical. RMT was was ultimately inspired by
and developed to solve complex, many-body problems in statistical
quantum mechanics (Guhr & Mehta, 1998). It was only recently with the
explosion of big data and more powerful computational capacities that
RMT has started to become relevant to other complex systems (Guhr &
Mehta, 1998).

## Interpretations
"The successful application of GRMT to
data shows that these data fluctuate stochastically in a manner
consistent with GRMT. Since GRMT can be derived from a maximum entropy
principle [18], this statement is tantamount to saying that the data
under study carry no information content" (Guhr & Mehta, 1998, p.202)

### RMT-Deviating Eigenvalues
"For complex quantum systems, RMT predictions represent an average
over all possible interactions [8–10]. Deviations from the universal
predictions of RMT identify system specific, non- random properties of
the system under consideration, provid- ing clues about the underlying
interactions [11–13]. Recent studies [14,15] applying RMT methods to
analyze the properties of C show that ~98% of the eigenvalues of C
agree with RMT predictions, suggesting a considerable de- gree of
randomness in the measured cross correlations. It is also found that
there are deviations from RMT predictions for ~2% of the largest
eigenvalues. These results prompt the following questions:" (Plerou et
al., 2002)

This seems to be a promising way to distinguish task from rest.

## Limitations
"It is clear that *RMT cannot ever reproduce a given data set in its
full detail*. It can only yield the distribution function of, and the
correlations between, the data points. An example is the nuclear level
spectrum at neutron threshold, the first case ever studied in a
statistically significant fashion. The actual sequence of levels
observed experimentally remains inaccessible to RMT predictions, while
the distribution of spacings and the correlations between levels can
be predicted. The same holds for the stochastic fluctuations of
nuclear cross sections, for universal conductance fluctu- ations of
mesoscopic probes, and for all other applications of RMT." (Guhr &
Mehta, 1998, p.202)

### Practical / Computational Issues

### Issues Related to fMRI