### 2021-09-08 paper/latex/rmt.tex

\input{preamble.tex}
\begin{document}

\maketitletwo[5]  % Optional argument is assignment number
%Keep a blank space between maketitletwo and \question[1]

\question[1]{Here is my first question}

YOUR SOLUTION HERE

\question[2]{Here is my second question}

YOUR SOLUTION HERE

\question[3]{What is the \Large{$\int_0^2 x^2 \, dx $}\normalsize{. Show all steps}}

\begin{align*}
\int_0^2 x^2 &= \left. \frac{x^3}{3} \right|_0^2 \\
                &= \frac{2^3}{3}-\frac{0^3}{3}\\
                &= \frac{8}{3}
\end{align*}

\end{document}