### 2022-10-06 paper/bookdown/rmt.Rmd
update

---
title: "Random Matrix Theory Tools for the Analysis of Functional Magnetic-Resonance Imaging Examinations"
fontsize: 10pt
urlcolor: blue
linkcolor: blue
# mainfont: "Garamond Premier Pro"
# biblio-style: "apalike"
bibliography: "rmt.bib"  # exported from Zotero with better BibLaTeX extension
link-citations: true
output:
    # bookdown::pdf_document2:
    bookdown::pdf_book:
        latex_engine: xelatex
        toc: false
        # toc_depth: 2
        number_sections: false
        highlight: tango
        # https://bookdown.org/yihui/rmarkdown-cookbook/figure-placement.html
        # unicode-math needed for bold italic math
        # (https://tex.stackexchange.com/questions/431013/error-extended-mathchar-used-as-mathchar-when-using-bm)
        extra_dependencies: ["float", "unicode-math"]
        mathjax: default
        includes:
            in_header: preamble.tex
        split_bib: true
        fig_width: 6
        fig_height: 4
        link-citations: true
citation-style: ieee.csl # https://www.zotero.org/styles
---

# Abstract

Previous studies have investigated the potential of using analytic techniques from Random Matrix
Theory (RMT) to investigate magnetic resonance imaging (MRI) data. We assess the potential
application of RMT-based features for the analysis of functional MRI (fMRI) across diverse datasets.
As novel contributions, we (1) assess the potential for RMT-inspired, whole-brain features extracted
from voxel-wise functional connectivity to contain information useful for classifying between
various psychological processes, (2) assess these features’ predictive—rather than
explanatory—value, (3) investigate the effect of varying RMT analysis methods on the robustness of
study findings, and (4) make general-purpose code publicly available for users to extract these
features from a wide variety of data. We find preliminary evidence suggesting that RMT-inspired
features may have unique potential in analyses of fMRI functional connectivity.

## Index Terms / Keywords

feature extraction, functional connectivity, functional magnetic resonance imaging, random matrix
theory, voxelwise connectivity

## Funding Info

Manuscript submitted January 20, 2021. This work was supported by the  Natural Science and
Engineering Research Council of Canada's Canada Research Chair grant (grant number 231266) to JL,
Natural Science and Engineering Research Council of Canada Discovery Grant to JL, a Canada
Foundation for Innovation and Nova Scotia Research and Innovation Trust infrastructure grant
(R0176004) to JL, a St. Francis Xavier University research startup grant to JL (grant number
R0168020), and a St. Francis Xavier University UCR grant to JL.

## Author Info

Derek Berger is with the department of Computer Science, St. Francis Xavier University, Antigonish
NS B2G 2W5, Canada (e-mail: dberger@stfx.ca).

Gurpreet Matharoo is with ACENET, and the Department of Physics, St. Francis Xavier University,
Antigonish NS B2G 2W5, Canada (e-mail: gmatharo@stfx.ca).

Jacob Levman is with the Department of Computer Science, St. Francis Xavier University, Antigonish
NS B2G 2W5, Canada (e-mail: jlevman@stfx.ca).

\newpage

# Introduction

In functional magnetic resonance imaging (fMRI), changes in the blood-oxygenation-level-dependent
(BOLD) signals are related to neural activity. It is common to investigate statistical relationships
between the BOLD signals through functional connectivity analyses, where correlations between
collections of these signals are examined to infer connections between different voxels or regions
of interest (ROIs) within the brain.

Whether in the presence of experimental stimuli, or the relative absence, as in a resting state,
complex functional connectivity networks are ubiquitous [1]–[4]. This complexity suggests fMRI is a
candidate to be studied using Random Matrix Theory (RMT), a set of mathematical tools originally
developed some 50 years ago to solve complex problems in nuclear physics [5], [6]. In various
physical systems where nuclei can have different possible states, precisely modeling the entirety of
the interactions between nuclei is often intractable. For such large, complex systems, the
eigenvalues (or spectra) efficiently summarize the totality of the interactions between components
of the system, and RMT describes the expected behaviour of such eigenvalues. By analysing the
spectrum statistically, complex systems can be investigated by comparison to the universal
properties predicted by RMT.

These universal properties have been observed in diverse phenomena. In small-scale systems, RMT
universalities have been observed in quantum chaotic systems, complex nuclei, atoms, molecules and
disordered mesoscopic systems [5]–[11], and at larger scales, RMT has been applied to atmospheric
physics [12], stock cross-correlations [13], social networks [14], random networks [15],
network-formation in liquids [16], [17] and amorphous clusters [18]–[21]. Within biological systems,
RMT has also been used to successfully model aspects of  amino acid functional relationships [22],
synchrony in epileptic seizures [23], and in protein-protein interactions both in different species
[24] and breast cancer [25].

In recent years, RMT analyses have been applied to study brain functioning.  The earliest  study
demonstrated that spectra of the correlations between electroencephalographic (EEG) signals closely
resemble those of the Gaussian Orthogonal Ensemble (GOE) [26].  Somewhat more recently, RMT was used
to evaluate the quality of whole brain features extracted from fMRI data [27], [28]. RMT has also
been used in diffusion MRI to aid in the selection of the number of components to employ in
principal-component reduction analysis and denoising [29]–[31].

Finally, RMT has been used in ROI-based fMRI functional-connectivity studies to investigate
differences between rest and task states [32], between subjects with and without ADHD [33], and
between pain and non-pain states [34]. Across these three studies, the spectra of resting or
low-attention states exhibited properties closest to the GOE.

These findings suggest that certain aspects of psychological processes might be characterized, in
part, by features computed from the eigenvalues of fMRI correlation matrices, and that these
features might vary in an interpretable manner across psychological processes. If this is the case,
RMT could have unique potential in characterizing the functioning of the human brain.

Here, we use a novel, voxel-based approach, and expand the applications of RMT to analyze fMRI scans
from  diverse datasets1. We extract RMT-inspired whole-brain features from voxelwise functional
connectivity data, and assess the predictive value of these features. We also examine in detail the
robustness of the RMT methodology to various common analytic choices.

# Materials and Methods

## Datasets

As our primary intention is to examine the potential utility of RMT-derived features across a wide
variety of data, and not to decisively put forward or refute any specific hypotheses relating RMT to
psychological processes, we examine a wide selection of fMRI datasets available on the OpenNeuro
platform [36].  This allows for a broad assessment of the potential of voxelwise RMT features.

However, since previous studies employing RMT for feature extraction [33], [34] have partially
interpreted their findings with respect to attentional processes, we have elected to use datasets
that involve attention in some way. This includes measures of attention or vigilance, or comparisons
between a condition involving more attention and focus (e.g. a task state) or conditions which
should involve considerably less attention or focus (e.g. resting-state).

```{r , echo=FALSE, results='asis'}
  cat(' Table: (\\#tab:scan-params) Summary of Scan Parameters. ID = Identifier for paper. FOV = Field of View. TR = Time of Repetition (seconds). Time = total duration (minutes) of each scan. Dimensions listed as M × N × P, indicate P slices with dimensions M × N.

  |   ID   |     FOV (mm) | Dimensions | Voxel Size (mm) |  TR  | Volumes | Time (m) |
  |:------:|:------------:|:-----------:|:---------------:|:----:|:-------:|:--------:|
  | PSYCH  | 192          | 128×128×70 | 1.5×1.5×1.5     | 3.0  |     300 |  15     |
  | OSTEO  | -            | 64×64×36   | 3.4×3.4×3.0     | 2.5  |     300 |  12.5   |
  | REFLECT| 210          | 72×72×46   | 3.0×3.0×3.0     | 3.0  |     204 |  10.2   |
  |        | 240          | 64×64×33   | 3.8×3.8×3.8     | 2.0  |     260 |  8.7    |
  | PARK   | 240×240×129  | 80×80×43   | 3.0×3.0×3.0     | 2.4  |     149 |  5.96   |
  | LEARN  |  -           | 64×64×36   | 3.0×3.0×3.0     | 2.0  |     195 |  6.5    |')

```

## A Simple Linear MLP

A citation to [@abueleninSpectralUnfoldingChaotic2018a;
@abul-magdModellingHighwaytrafficHeadway2007; @alakorkkoEffectsSpatialSmoothing2017].

As a concrete example, imagine we are build a simple ANN from $\mathbbm{R}$ to $\mathbbm{R}^2$ by
way of two larger intermediate layers. We might have for example

```python
ann = Sequential([
    Linear(1, 2),
    Linear(2, 1),
])
```

which learns the weights

$$
\begin{aligned}
\mathbf{W}_1 &= \begin{bmatrix} w_1 & w_2 \end{bmatrix} & \text{layer 1: matrix of shape (1, 2), output shape (1, 2)} \\
\mathbf{W}_2 &= \begin{bmatrix} w_3 \\ w_4 \end{bmatrix}& \text{layer 2: matrix of shape (2, 1), output shape (1, 1)} \\
\end{aligned}
$$

Assuming ReLUs and no biases, the entire equation for this network is
$y = \mathbf{W}_2 \cdot \text{ReLU}(\mathbf{W}_1 x) + b$ (final layer should always have a bias), that is:

$$
\begin{aligned}
\mathbf{W}_1 x & = \begin{bmatrix} w_1 \cdot x & \;\; w_2 \cdot x \end{bmatrix} \\
             y & = \begin{bmatrix}  w_3 \cdot \text{ReLU}(w_1 \cdot x) + w_4 \cdot \text{ReLU}(w_2 \cdot x) + b \end{bmatrix} \\
\end{aligned}
$$

This in fact reduces to:

$$
y = \begin{cases}
w_1 w_3 x + w_2 w_4 x + b  & \text{ if }  sgn(x) =   sgn(w_1) \text{ and } sgn(x)   = sgn(w_2) \\
w_1 w_3 x + b              & \text{ if }  sgn(x) =   sgn(w_1) \text{ and } sgn(x) \ne sgn(w_2) \\
w_2 w_4 x + b              & \text{ if }  sgn(x) \ne sgn(w_1) \text{ and } sgn(x)   = sgn(w_2) \\
b                          & x = 0 \\
\end{cases}
$$

In particular, without loss of generality we can take $w_1 < 0$ and $w_2 > 0$ and reduce this to:

$$
y = \begin{cases}
w_2 w_4 x + b  & \text{ if }  x \ge 0 \\
w_1 w_3 x + b  & \text{ if }  x  < 0 \\
\end{cases}
$$

# References