### 2022-11-12 paper/paper/SPIE/article.tex
[WIP]

\documentclass[12pt]{spieman}  % 12pt font required by SPIE;
%\documentclass[a4paper,12pt]{spieman}  % use this instead for A4 paper
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{graphicx}
\graphicspath{ {./figures/} }
\usepackage{setspace}
\usepackage{tocloft}
\usepackage{doi}
\usepackage{csquotes}
\MakeOuterQuote{"}
\usepackage{xr}
\externaldocument[S-]{supplementary}

\newcommand{\code}[1]{\small\texttt{#1}\normalsize}
\newcommand{\tcode}[1]{\footnotesize\texttt{#1}\normalsize}

\title{Random Matrix Theory Tools for the Analysis of Functional Magnetic Resonance Imaging Examinations}

\author[a]{Derek Berger}
\author[a,*]{Jacob Levman}
\author[b]{Gurpreet M. Matharoo}
\affil[a]{St. Francis Xavier University, Department of Computer Science, 4130 University Avenue, Antigonish, Canada, B2G 2W5}
\affil[b]{St. Francis Xavier University, ACENET, 4130 University Avenue, Antigonish, Canada, B2G 2W5}

\renewcommand{\cftdotsep}{\cftnodots}
\cftpagenumbersoff{figure}
\cftpagenumbersoff{table}
\begin{document}
\maketitle

% \begin{abstract} This document shows the required format and appearance of a
% manuscript prepared for SPIE journals. It is prepared using LaTeX2e with the
% class file \texttt{spieman.cls}. Please note that the following journals
% require the use of structured abstracts in manuscript submissions:
% \textit{Neurophotonics}, the \textit{Journal of Biomedical Optics}, and the
% \textit{Journal of Medical Imaging}. Structured abstracts are encouraged for
% the \textit{Journal of Micro/Nanolithography, MEMS, and MOEMS}. Guidelines
% are available on the journal website. Whether structured or single-paragraph,
% the abstract should be a summary of the paper and not an introduction.
% Because the abstract may be used in abstracting and indexing databases, it
% should be self-contained (i.e., no numerical references) and substantive in
% nature, presenting concisely the objectives, methodology used, results
% obtained, and their significance. A list of up to six keywords should
% immediately follow. \end{abstract}

% See
% https://www.spiedigitallibrary.org/journals/journal-of-medical-imaging/volume-7/issue-01/010101/Integrating-structured-abstracts-in-the-Journal-of-Medical-Imaging/10.1117/1.JMI.7.1.010101.full
% for info on the structured abstract
%
% Purpose: This section presents the significance and aims, stating the broad
% impact and the rationale or motivation of the work, including potentially
% some background.
%
% Approach: This section briefly describes the materials and methods, including
% the statistical analyses, used in the research.
%
% Results: This section provides a core summary of the findings from the work,
% including study numbers, quantitative analyses, or discoveries.
%
% Conclusions: This section summarizes and interprets the approach and findings
% from the work, stating also the importance or impact of the findings.

% Purpose: Data-intensive modeling could provide insight on the broad
% variability in outcomes in spine surgery. Previous studies were limited to
% analysis of demographic and clinical characteristics. We report an analytic
% framework called “SpineCloud” that incorporates quantitative features
% extracted from perioperative images to predict spine surgery outcome.

% Approach: A retrospective study was conducted in which patient demographics,
% imaging, and outcome data were collected. Image features were automatically
% computed from perioperative CT. Postoperative 3- and 12-month functional and
% pain outcomes were analyzed in terms of improvement relative to the
% preoperative state. A boosted decision tree classifier was trained to predict
% outcome using demographic and image features as predictor variables.
% Predictions were computed based on SpineCloud and conventional demographic
% models, and features associated with poor outcome were identified from
% weighting terms evident in the boosted tree.

% Results Neither approach was predictive of 3- or 12-month outcomes based on
% preoperative data alone in the current, preliminary study. However,
% SpineCloud predictions incorporating image features obtained during and
% immediately following surgery (i.e., intraoperative and immediate
% postoperative images) exhibited significant improvement in area under the
% receiver operating characteristic (AUC): AUC=0.72 (CI95=0.59 to 0.83) at 3
% months and AUC=0.69 (CI95=0.55 to 0.82) at 12 months.

% Conclusions Predictive modeling of lumbar spine surgery outcomes was improved
% by incorporation of image-based features compared to analysis based on
% conventional demographic data. The SpineCloud framework could improve
% understanding of factors underlying outcome variability and warrants further
% investigation and validation in a larger patient cohort.
\begin{abstract}

\section*{Purpose}
Random matrix theory (RMT) is an increasingly useful tool for understanding
large, complex systems. Prior studies have examined functional Magnetic
Resonance Imaging (fMRI) scans using tools from RMT, with some success.
However, RMT computations are highly sensitive to a number of analytic choices,
and the robustness of findings involving RMT remain in question. We
systematically investigate the usefulness of RMT on a wide variety of fMRI
datasets using a rigorous predictive framework.

\section*{Approach}
We develop open-source software to efficiently compute RMT features from fMRI
images, and examine the cross-validated predictive potential of eigenvalue and
RMT-based features (``eigenfeatures'') with classic machine-learning
classifiers. We systematically vary pre-processing extent, normalization
procedures, RMT unfolding procedures, and feature selection, and compare the
impact of these analytic choices on the distributions of cross-validated
prediction performance for each combination of dataset binary classification
task, classifier, and feature. To deal with class imbalance, we use the
area under the receiver operating characteristic (AUROC) as the main
performance metric.

\section*{Results}
Across all classification tasks and analytic choices, we find RMT- and
eigenvalue-based "eigenfeatures" to have predictive utility more often than
not (82.4\% of median AUROCs \(>\) 0.5; median AUROC range across
classification tasks 0.47 - 0.64). Simple baseline reductions on source
timeseries, by contrast, were less useful (58.8\% of median AUROCs \(>\) 0.5,
median AUROC range across classification tasks 0.42 - 0.62). Additionally,
eigenfeature AUROC distributions were overall more right-tailed than
baseline features, suggesting greater predictive potential. However, performance
distributions were wide, and often significantly affected by analytic choices.

\section*{Conclusions}
Eigenfeatures clearly have potential for understanding and reducing fMRI
functional connectivity. However, the utility of these features is strongly
dependent on analytic decisions, suggesting considerable caution when
interpreting past and future studies applying RMT to fMRI.

% This document shows the required format and appearance of a manuscript
% prepared for SPIE journals. It is prepared using LaTeX2e with the class file
% \texttt{spieman.cls}. Please note that the following journals require the use
% of structured abstracts in manuscript submissions: \textit{Neurophotonics},
% the \textit{Journal of Biomedical Optics}, and the \textit{Journal of Medical
% Imaging}. Structured abstracts are encouraged for the \textit{Journal of
% Micro/Nanolithography, MEMS, and MOEMS}. Guidelines are available on the
% journal website. Whether structured or single-paragraph, the abstract should
% be a summary of the paper and not an introduction. Because the abstract may
% be used in abstracting and indexing databases, it should be self-contained
% (i.e., no numerical references) and substantive in nature, presenting
% concisely the objectives, methodology used, results obtained, and their
% significance. A list of up to six keywords should immediately follow.
\end{abstract}

% Include a list of up to six keywords after the abstract
\keywords{random matrix, spectral rigidity, level number variance, fMRI, classification, machine-learning}

% Include email contact information for corresponding author
{\noindent \footnotesize\textbf{*}Jacob Levman,  \linkable{jlevman@stfx.ca} }

\begin{spacing}{2}   % use double spacing for rest of manuscript

\section{Introduction}
\label{sect:intro}  % \label{} allows reference to this section
% This document shows the format and appearance of a manuscript prepared for
% submission to an SPIE journal. Note that this template is only intended to be
% used as a guideline for author convenience. It is designed for optimum
% clarity and ease of reading for editors and reviewers, but the template does
% not reflect the final page layout of a published journal paper. Accepted
% papers are professionally typeset in XML according to the layout and design
% of the journal.

At its most rudimentary, RMT describes the expected behavior of the
eigenvalues—also often called levels—of a number of classes of random
matrices\cite{andersonIntroductionRandomMatrices2010,
akemannOxfordHandbookRandom2011, mehtaRandomMatrices2004}. A random matrix is,
simplifying somewhat, a matrix in which the entries are drawn from
independently and identically distributed (\textit{iid}) distributions. Though
first developed to describe the fluctuation of nuclei energy levels in quantum
physics \cite{mehtaRandomMatrices2004, guhrRandommatrixTheoriesQuantum1998a},
RMT has been shown have extremely broad potential.

In small-scale physical systems, RMT universalities have been observed in
quantum chaotic systems, complex nuclei, atoms, molecules and disordered
mesoscopic systems
\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004,brodyRandommatrixPhysicsSpectrum1981,beenakkerRandommatrixTheoryQuantum1997,bohigasHigherOrderCorrelationsSpectra1985,wintgenLevelStatisticsQuantized1988,pandeySkewOrthogonalPolynomialsUniversality2001},
and at larger scales, RMT has been applied to atmospheric physics
\cite{santhanamStatisticsAtmosphericCorrelations2001}, stock cross-correlations
\cite{plerouRandomMatrixApproach2002} , social networks
\cite{jalanUncoveringRandomnessSuccess2014}, random networks
\cite{bandyopadhyayUniversalityComplexNetworks2007}, network-formation in
liquids
\cite{sastrySpectralStatisticsInstantaneous2001,matharooSpectralStatisticsQuenched2009}
and amorphous clusters
\cite{sarkarUniversalityVibrationalSpectra2004,matharooVibrationalSpectraAmorphous2005,matharooUniversalityVibrationalSpectra2008}.
Within biological systems, RMT has also been used to successfully model aspects
of  amino acid functional relationships
\cite{bhadolaTargetingFunctionalMotifs2016}, synchrony in epileptic seizures
\cite{osorioPhasesynchronizationRandommatrixBased2011}, and in protein-protein
interactions both in different species
\cite{agrawalQuantifyingRandomnessProtein2014} and breast cancer
\cite{raiRandomnessPreservedPatterns2015}. RMT has also been used to guide
statistical decisions in principal components
analyses\cite{franklinParallelAnalysisMethod1995,
veraartDenoisingDiffusionMRI2016, ulfarssonDimensionEstimationNoisy2008}, and,
even more recently, has provided insights into the behaviors of deep
neural networks\cite{martinImplicitSelfRegularizationDeep2021,
martinPredictingTrendsQuality2021}.

RMT likely has the most real-world explanatory potential when dealing with a
complex system of many (hundreds or more\cite{mehtaRandomMatrices2004})
interacting components. If such a system has a matrix representation, and the
eigenvalues of this representation have a distribution similar to one predicted
by RMT, it suggests the system is either highly random or chaotic. By contrast,
if the observed spectra deviate significantly from RMT-predicted spectra, this
suggests otherwise. A number of studies have used RMT to make such
interpretations and comparisons between
systems\cite{santhanamStatisticsAtmosphericCorrelations2001,
jalanUncoveringRandomnessSuccess2014,
bandyopadhyayUniversalityComplexNetworks2007,
agrawalQuantifyingRandomnessProtein2014, raiRandomnessPreservedPatterns2015,
sebaRandomMatrixAnalysis2003, wangSpectralPropertiesTemporal2015,
wangRandomMatrixTheory2016, matharooSpontaneousBackpainAlters2020}.

\subsection{RMT and Neurobiological Signals}

In the human brain, each neuron, collection of neurons, or region of interest
(ROI) is a potentially-interacting component in a complex system. RMT
may have potential in describing the totality of these interactions, provided
that measurements of functioning can be obtained with sufficient spatial and
temporal resolution to speak to some neurobiological or neuropsychological
phenomenon of interest.

For an imaging modality like fMRI, where changes in the
blood-oxygenation-level-dependent (BOLD) signals are related to neural
activity, RMT may be an ideal starting point, as each voxel time course (or
collection of such time sources - i.e. ROIs) can be considered an interacting
component of the system. Likewise, in functional connectivity
analyses, statistical relationships of the BOLD ROI time-courses are
investigated in the hope of gaining insights into brain functioning (see
\citenum{bastosTutorialReviewFunctional2016,
vandenheuvelExploringBrainNetwork2010} for reviews, but also
\citenum{nobleDecadeTestretestReliability2019} and
\citenum{lurieQuestionsControversiesStudy2020} for challenges facing functional
connectivity analyses). In this framework, each connection or correlation can
be considered a system component, and the eigenvalues of such a correlation
matrix can be examined from the perspective of RMT.

The earliest study taking this approach
demonstrated that spectra of the correlations between electroencephalographic
(EEG) signals closely resemble those of the Gaussian Orthogonal Ensemble (GOE)
\cite{sebaRandomMatrixAnalysis2003}. In fMRI, RMT has been used to
evaluate the quality of whole brain features extracted from fMRI data
\cite{voultsidouFeatureEvaluationFMRI2007,verganiRestingStateFMRI2019}, and in
diffusion MRI to aid in the selection of the number of
components to employ in principal-component reduction analysis and denoising
\cite{veraartDenoisingDiffusionMRI2016,verganiRestingStateFMRI2019,ulfarssonDimensionEstimationNoisy2008}.

RMT has also been used in ROI-based fMRI functional connectivity studies to
investigate differences between rest and task states
\cite{wangSpectralPropertiesTemporal2015}, between subjects with and without
attention-deficit hyperactive disorder (ADHD)
\cite{wangRandomMatrixTheory2016}, between pain and non-pain states
\cite{matharooSpontaneousBackpainAlters2020}, and between left-sided vs.
right-sided motor imagery \cite{guRandomMatrixTheory2020}. While no differences
were found in the latter study, across the first three studies, the spectra of
resting or low-attention states exhibited properties close to the GOE. These
findings suggest that certain aspects of psychological processes might be
characterized, in part, by features computed from the eigenvalues of fMRI
correlation matrices, and that these features might vary in an interpretable
manner across psychological processes. If this is the case, RMT could aid in
understanding the functioning of the human brain.

\subsection{Eigenvalue Features}

The basic insight of RMT is thus that eigenvalues alone may provide
interesting information about highly complex systems. However, real systems are
usually noisy, and involve a mixture of random and non-random components and
interactions. RMT is statistical in nature, describing only the expected
behavior of the spectra of \textit{iid} random matrices. To this end, a number
of summary statistics or ``spectral observables''
\cite{mehtaRandomMatrices2004, guhrRandommatrixTheoriesQuantum1998a} can be
computed from empirically observed spectra, with these spectral observables
sometimes being better suited for further analysis, or for the comparison of
empirical observations to theory. Two such summary statistics that have been
popular\cite{santhanamStatisticsAtmosphericCorrelations2001,
jalanUncoveringRandomnessSuccess2014, matharooSpontaneousBackpainAlters2020,
bandyopadhyayUniversalityComplexNetworks2007,
agrawalQuantifyingRandomnessProtein2014, raiRandomnessPreservedPatterns2015,
sebaRandomMatrixAnalysis2003,wangSpectralPropertiesTemporal2015,
wangRandomMatrixTheory2016} are the \textit{spectral rigidity} and
\textit{level number variance} (``rigidity'' and ``level variance'' for short;
details in Section \ref{sec:methods}).

However, from a predictive standpoint, these summary statistics may eliminate
predictively-useful information. If the basic RMT insight is that the
eigenvalues alone can provide understanding of a system, then those eigenvalues
(or other simple transformations of them) ought also to be predictively useful.
We examine a number of such features in this study, and refer to both
RMT-derived features like the rigidity and level variance, and non-RMT-derived
features collectively as \textit{eigenfeatures} in this study.

\subsection{Functional Connectivity Reduction}

The functional connectivity—the matrix of correlations of
various collections of voxels of an fMRI scan—is \textit{a priori} a useful
representation of the fMRI data. However, the full correlation matrix between
all voxels is itself often computationally infeasible to work with, and thus is
reduced in various ways prior to being used in analyses.

Typical reductions include the use of a ``seed'' voxel or collection of voxels
(see e.g. \citenum{joelRelationshipSeedbasedICAbased2011,
smithCharacterizingIndividualDifferences2014}): the mean signal over that ROI
is correlated with all other N ROIs of interest to generate a reduced
functional connectivity matrix (or image) with one correlation value at each
non-seed ROI. This sort of reduction reduces the functional connectivity to N
values, but necessarily biases the analysis to the seed ROI.

Likewise, one can work with ROI mean signal reductions, or independent
component analysis reductions\cite{joelRelationshipSeedbasedICAbased2011,
smithCharacterizingIndividualDifferences2014}, and work with the \(N \times N\)
matrix of the correlations of these reductions. This analysis is, \textit{a
priori}, sensitive to the choice of ROIs, and, in the case of anatomical
atlases / parcellations for ROIs, may also lack theoretical justification.

RMT suggests a potentially useful reduction in the form of the eigenvalues of
the voxelwise functional connectivity. An \(N \times t\) matrix of
\(N\) time series of length \(t\), and where \(t \ll N\) will have a symmetric
correlation (or covariance) matrix with \(t - 1\) non-zero, positive
real-valued eigenvalues. These eigenvalues can be computed highly
efficiently via transposition of the voxelwise correlation or covariance
matrix.

\subsection{Limitations of Previous Work}

Previous studies applying RMT to functional connectivity
data\cite{wangRandomMatrixTheory2016,
wangSpectralPropertiesTemporal2015a,matharooSpontaneousBackpainAlters2020,
guRandomMatrixTheory2020} took largely descriptive / explanatory approaches,
noting only differences in RMT across subgroups and/or conditions. For example
\citenum{wangRandomMatrixTheory2016} and \citenum{guRandomMatrixTheory2020} use
a Kolmogorov-Smirnov test, and \citenum{wangSpectralPropertiesTemporal2015a}
uses a t-test to provide evidence of differences in various RMT metrics.
However, statistically significant differences in a metric do not imply
practical significance or predictive utility
\cite{loWhySignificantVariables2015}, nor do they imply generalization or
replicability \cite{amrheinEarthFlat052017}. Cross-validation, by contrast,
attempts to more directly assess these properties
\cite{yarkoniChoosingPredictionExplanation2017}.

In addition, the previous papers do not provide publicly-accessible code to
reproduce results. The extraction of the spectral rigidity and level variance
is computationally-demanding and mathematically and algorithmically
non-trivial, and additionally requires an \textit{unfolding}
procedure\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}.
Unfolding is an exponential fitting procedure that involves a number of highly
subjective decisions regarding outliers and the flexibility of the fitting
function. These decisions are, unfortunately, often poorly or even entirely
undocumented, but are known to often dramatically impact RMT conclusions
\cite{abul-magdUnfoldingSpectrumChaotic2014,
abueleninSpectralUnfoldingChaotic2018, fossionRandommatrixSpectraTime2013,
abueleninEffectUnfoldingSpectral2012, moralesImprovedUnfoldingDetrending2011}.

On top of the flexibility in the implementation and application of RMT to
empirical data, there is also the flexibility introduced by the choices in the
complex preprocessing pipelines of fMRI\cite{parkerBenefitSliceTiming2019}.
These "researcher degrees of
freedom"\cite{simmonsFalsePositivePsychologyUndisclosed2011b} in combination
with the absence of reproducible code and data availability, and predominance
of descriptive and explanatory approaches, raise doubts about the basic robustness
and practical utility of RMT-based functional connectivity analyses.

What is missing is a rigorous, systematic, reproducible investigation of the
predictive value of RMT metrics across a wide variety of data, analytic
choices, RMT features, and preprocessing decisions, with RMT features being
compared to simpler alternative baseline predictors. The current study aims to
remedy this.

\section{Datasets}
\label{sec:datasets}

\subsection{Overview}

We selected fMRI data publicly available on the
\href{https://openneuro.org/}{OpenNeuro} platform (36). Selection criteria was
somewhat subjective, but we required (1), the dataset comprise 10+ human
subjects, (2) that all fMRI scans have the same spatial and temporal
resolutions, (3) that the fMRI images could be split into various classes for a
number of binary classification tasks, and (4), that it was reasonable to
classify each scan in its entirety, (i.e. neither task nor event timings
were needed to determine class membership, nor did any scan volumes need to be removed
to make the scan classifiable). This yielded seven datasets total.

\begin{table}[h!]
\caption{
    ID = Identifier for paper. TR = Time of Repetition (seconds).
    Time = total duration (minutes) of each scan. Dimensions listed as M x N x P,
    indicate P horizontal / axial slices each with dimensions M x N.
}
\label{table:1}
\small
\centering
\begin{tabular}{ l c c c c c c }
\hline
\textbf{Name}    & \textbf{Dimensions}  & \textbf{Voxel size (mm)} & \textbf{TR} & \textbf{Volumes} & \textbf{n\_scans} \\
\hline
Aging     & 74 × 74 × 32   & 3.0 × 3.0 × 4.0 & 2.0  & 300 & 62  \\
Bilingual    & 100 × 100 × 72 & 1.8 × 1.8 × 1.8 & 0.88 & 823 & 90  \\
Depress   & 112 × 112 × 25 & 2.0 × 2.0 × 5.0 & 2.5  & 100 & 72  \\
Learn     & 64 × 64 × 36   & 3.0 × 3.0 × 3.0 & 2.0  & 195 & 432 \\
Osteo     & 64 × 64 × 36   & 3.4 × 3.4 × 3.0 & 2.5  & 300 & 74  \\
Park      & 80 × 80 × 43   & 3.0 × 3.0 × 3.0 & 2.4  & 149 & 552 \\
Attention & 128 × 128 × 70 & 1.5 × 1.5 × 1.5 & 3.0  & 300 & 90  \\
\hline
\end{tabular}
\end{table}

In order not to inundate readers with dataset details, and because this is an
exploratory investigation of RMT which is not committed to any specific
theory\footnote{In fact, datasets and, later, the classification tasks, were
chosen without investigation into any findings or publications associated with
the data, both in order to reduce bias in our decisions, and because the
original authors analytic / statistical approaches and intentions have very
limited relevance to our whole-brain, voxelwise, multiverse predictive
approach.}, we refer the reader to the original publications and/or data
releases when greater detail is needed, and highlight only the most basic
aspects of each dataset here.  Dataset scan parameters and acquisition details
are summarized in Table \ref{table:1}, and sample and subgroup sizes are
summarized in Table \ref{table:2}.

\begin{table}[h!]
\caption{
    ID = Identifier for paper. Subgroup = name of subgroup. Subjects = Number of subjects.
    Task = fMRI task. ANT = Attention Network Task \cite{fanActivationAttentionalNetworks2005}
}
\label{table:2}
\small
\centering
\begin{tabular}{ l r c c c }
\hline
\textbf{ID}  & \textbf{Subgroup}  & \textbf{Subjects}  & \textbf{Task} & \textbf{Scans per Subject} \\
\hline
Aging   &  older             & 28      & resting-state        & 1 \\
        &  younger           & 34      & resting-state        & 1 \\
        &                    &         &                      &   \\
Bilingual  &  bilingual          & 59      & resting-state    & 1 \\
        &  monolingual        & 33      & resting-state       & 1 \\
        &                    &         &                      &   \\
Depress &  depression         & 51      & resting-state       & 1 \\
        &  control            & 21      & resting-state       & 1 \\
        &                    &         &                      &   \\
Learn   &  task              & 24      & learn image sketches & 16\\
        &  rest              & 24      & resting-state        & 2 \\
        &                    &         &                      &   \\
Osteo   &  duloxetine        & 19      & resting-state        & 1 \\
        &  pain              & 37      & resting-state        & 1 \\
        &  nopain            & 20      & resting-state        & 1 \\
        &                    &         &                      &   \\
Park    &  Parkinson's       & 25      & ANT                  & 12\\
        &  control           & 21      & ANT                  & 12\\
        &                    &         &                      &   \\
Attention&  vigilant           & 11     & resting-state       & 2 \\
        &  nonvigilant         & 11     & resting-state        & 2 \\
        &  trait-attentive     & 11    & resting-state        & 2 \\
        &  trait-non-attentive & 11    & resting-state        & 2 \\
        &  task-attentive     & 11     & resting-state        & 2 \\
        &  task-nonattentive  & 11     & resting-state        & 2 \\
\hline
\end{tabular}
\end{table}

\subsection{Datasets and Classification Tasks}

The \textit{Aging} dataset\cite{wahlheimIntrinsicFunctionalConnectivity2021}
included rs-fMRI data for 34 subjects with a mean age of 22 years (range: 18-32
years), and 28 subjects with a mean age of 70 years (range: 61-80 years). For
our analysis, we make the binary classification task the prediction of subject
age-group membership, i.e. \code{younger v older}.

% The final sample included 34 younger adults (1832 years old, Mage = 22.21, SD
% = 3.65; 20 female) and 28 older adults (61-80 years old, Mage = 69.82, SD =
% 5.64; 20 female). Table 1 shows that younger and older adults had comparable
% education and working memory capacity, the latter measured by forward and
% backward digit span (48). Older adults had higher vocabulary scores (49) and
% slower processing speed (50) than younger adults. Resting-state data
% comprised 300 measurements collected over one 10-minute run. This scan
% duration produces reliable functional connectivity estimates within and
% across participants (57). Participants were instructed to remain still and
% awake with their eyes open. No stimuli appeared during this scan, and the
% display was black. Functional scans were collected using an echo-planar image
% sequence sensitive to blood-oxygen-level-dependent (BOLD) contrast (T2*; 32
% slices with 4.0 mm thickness and no skip, TE = 30 ms, TR = 2000 ms, flip
% angle = 70, FOV = 220 mm, matrix size = 74 × 74 × 32 voxels, A/P phase
% encoding direction). Slices were collected in a descending order that covered
% the entire cortex and partial cerebellum. At the beginning of the
% resting-state scan, the scanner acquired and discarded two dummy scans.

% \subsubsection{Bilinguality Dataset}

The \textit{Bilinguality} dataset\cite{ds001747:1.0.0,
goldcarrieelizabethExploringRestingState2018} examined English and
Spanish-speaking monolinguals and multilinguals during a prolonged resting
state. The original study grouped participants into three subgroups: early vs
late bilinguals vs monolingual
controls\cite{goldcarrieelizabethExploringRestingState2018}. For our analysis,
we make the binary classification task \code{monolingual v
bilingual}.

% Resting state fMRI scans were performed using a T2*-weighted EPI sequence with the following
% parameters: TR=875ms; TE=43.6ms; slice thickness=1.8mm; acquisition matrix=100 × 100; flip
% angle=55°; number of slices=72; field of view=180 × 180mm; voxel size=1.8 × 1.8 × 1.8mm;
% multi-band factor=8; volumes per run=823 (total scan time=12min). EPI scans were oriented
% parallel to the long axis of the hippocampus.

% \subsubsection{Depression Dataset}

The \textit{Depression} dataset\cite{bezmaternykhddRestingStateClosed2021,
bezmaternykhBrainNetworksConnectivity2021} includes rs-fMRI scans from
non-depressed controls and mildly- or moderately-depressed subjects. In our analysis,
the binary classification task is \code{depress v control}.

% The intergroup study involved 21 healthy volunteers received a compensation
% for a participation and 51 participants featuring mild depression (F32.0),
% moderate depression (F32.1), or dysthymia (F34.1, single patient) (see Table
% 1). The fMRI study was carried out in the International Tomography Center,
% Novosibirsk, using a 3 T Ingenia scanner (Philips). Functional T2∗-weighted
% Ssh echo planar imaging scans were acquired using the following parameters:
% voxel size 2×2×5mm, repetition time/echo time = 2500/35 ms, and fat
% suppression mode. The reference anatomical image was obtained by the T1W 3D
% turbo field echo method with a voxel size of 1×1×1mm. The instruction for
% participants was to lie still with eyes closed for 6 minutes.

% \subsubsection{Learning Dataset}

The \textit{Learning} dataset\cite{schapiroHumanHippocampalReplay2018,
schapiroHumanHippocampalReplay2020} has both rs-fMRI and task fMRI scans
available for all subjects, and \code{task v rest} is chosen as the binary
classification task for our analysis.  The task is complex and difficult to
summarize here adequately, but involved multiple phases where subjects were
presented with various related abstract images, and then later tested for their
memory of certain aspects of the learned images. Thus, the classification task
might also be considered learning vs. rest.

% \subsubsection{Osteoarthritis Dataset}
The \textit{Osteo} dataset\cite{tetreaultBrainConnectivityPredicts2016}
includes whole-brain rs-fMRI scans of healthy, pain-free controls
(\code{nopain} condition), and individuals with knee osteoarthritis.
Osteoarthritic patients were treated for two weeks with either placebo
(\code{pain} condition) or duloxetine (\code{duloxetine} condition).
We use the three binary classification tasks \code{nopain v duloxetine},
\code{nopain v pain}, and \code{pain v duloxetine} for our analysis.

% \subsubsection{Parkinson's Dataset}
The \textit{Park} dataset\cite{madhyasthaDynamicConnectivityRest2015} includes
subjects with non-demented Parkinson's Disease (PD) and
healthy controls performed a number of repetitions of the Attention Network
Task (ANT\cite{fanActivationAttentionalNetworks2005}) during scans.
The classification task for this dataset is Parkinson's versus controls, i.e.
\code{park v ctrl}.

% \subsubsection{Attention Dataset}

The \textit{Attention} dataset\cite{gorgolewskiHighResolution7Tesla2015} is a
high-resolution rs-fMRI dataset including a battery of psychological measures
for each subject. Since previous studies employing RMT sometimes interpreted
their findings with respect to attentional processes
\cite{wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020}, we
divided subjects into various high vs. low attention binary classification
tasks based on median splits of subsets of the metrics available in this study.

The \code{viglant v nonvigilant} task was formed based on self-report
questionnaire items involving
"vigilance"\cite{gorgolewskiCorrespondenceIndividualDifferences2014}. The
\code{trait\_attend v trait\_nonattend} task was formed using PANAS-X
\cite{watsonPANASXManualPositive1994} items related to self-reported
wakefulness and attention over the past weeks. Finally, the
\code{task\_attend v task\_nonattend} classification task was constructed
using the scores of the Conjunctive Continuous Performance Task
(CCPT\cite{shalevConjunctiveContinuousPerformance2011}). This is a behavioral
performance task which requires the subject to quickly and selectively respond
to only certain visual stimuli. Details allowing exact reproduction of these
splits are available in this paper's
\href{https://github.com/DM-Berger/random-matrix-fmri/blob/master/code/rmt/preprocess/unify\_attention\_data.py}{source
code}.

\section{Methods}
\label{sec:methods}

\subsection{fMRI Preprocessing}
\label{sec:preproc}

We limited preprocessing to a simple pipeline of, in order: (1) brain
extraction, (2) slice timing correction, (3), motion correction, and (4),
registration to the MNI152
\cite{fonovUnbiasedAverageAgeappropriate2011,fonovUnbiasedNonlinearAverage2009}
2mm isotropic, asymmetric template version C made available through
\href{https://www.templateflow.org/}{TemplateFlow}
\cite{ciricTemplateFlowFAIRsharingMultiscale2021}.

Brain extraction was performed first with BET
\cite{smithFastRobustAutomated2002}, and then with with
ANTs\cite{avantsReproducibleEvaluationANTs2011} to clean up residue left behind
by BET. BET results were visually inspected for each image in each dataset, and
BET fractional anisotropy values were modified to ensure brain extraction was
acceptable. Motion-correction was performed with FSL's MCFLIRT
\cite{jenkinsonImprovedOptimizationRobust2002}, and slicetime correction also
with FSL's \code{slicetimer}\cite{jenkinsonFSL2012} tool. Finally, functional
images were registered directly to the MNI152 template via
ANTs\cite{avantsReproducibleEvaluationANTs2011}.

We limited preprocessing methods to these steps sometimes because data required
for other typical preprocessing steps (e.g. field intensity maps, physiological
measurements) was missing, but also because all pipeline intermediates were
saved to later compare the effect of increasing degrees of preprocessing on
feature predictive utility (see Section \ref{sec:multiverse}), and it was
important to limit the number of preprocessing steps to prevent excessive
computational costs and comparisons.

\subsection{Feature Extraction}

For each preprocessing degree and fMRI image, features were extracted from all
\(N\) non-constant brain voxels, where \(N\) may differ for each image. After
this process, each scan yields an \(N \times T\) matrix \(\mathbf{M}\), where
\(T\) is the number of volumes acquired. All features summarize this matrix
of brain voxels \(\mathbf{M}\) in some manner.

\begin{table}[h!]
\caption{
\label{tab:features}
Feature groupings for summarization. eigs = non-RMT eigenfeatures and their combinations with
other non-RMT eigenfeatures. rmt = RMT eigenfeatures and their combinations. tseries = baseline
timeseries reductions.\\
}
\small
\centering
\begin{tabular}{lllr}
\hline
\textbf{Coarse Grouping} & \textbf{Fine Grouping} & \textbf{Feature id}   \\
\hline
eigs           & eigs          & \tcode{eigs}                           \\
               & eigs max      & \tcode{eigsminmax10}                   \\
               &               & \tcode{eigsminmax20}                   \\
               &               & \tcode{eigsminmax5}                    \\
               & eigs middle   & \tcode{eigsmiddle10}                   \\
               &               & \tcode{eigsmiddle20}                   \\
               &               & \tcode{eigsmiddle40}                   \\
               & eigs smooth   & \tcode{eigs + eigs\_smooth}            \\
               &               & \tcode{eigs + savgol}                  \\
               &               & \tcode{eigs\_savgol}                   \\
               &               & \tcode{eigs\_smooth}                   \\
               &               &                                        \\
rmt            & rmt + eigs    & \tcode{eigs + levelvar}                \\
               &               & \tcode{eigs + rigidity}                \\
               &               & \tcode{eigs + rigidity + levelvar}     \\
               &               & \tcode{eigs + unfolded}                \\
               &               & \tcode{eigs + unfolded + levelvar}     \\
               &               & \tcode{eigs + unfolded + rigidity}     \\
               & rmt only      & \tcode{levelvar}                       \\
               &               & \tcode{rigidity}                       \\
               &               & \tcode{rigidity + levelvar}            \\
               &               & \tcode{unfolded}                       \\
               &               & \tcode{unfolded + levelvar}            \\
               &               & \tcode{unfolded + rigidity}            \\
               &               & \tcode{unfolded + rigidity + levelvar} \\
               &               &                                        \\
tseries        & location      & \tcode{T-max}                          \\
               &               & \tcode{T-mean}                         \\
               &               & \tcode{T-med}                          \\
               &               & \tcode{T-min}                          \\
               &               & \tcode{T-p05}                          \\
               &               & \tcode{T-p95}                          \\
               & scale         & \tcode{T-iqr}                          \\
               &               & \tcode{T-rng}                          \\
               &               & \tcode{T-rrng}                         \\
               &               & \tcode{T-std}                          \\
\hline
\end{tabular}
\end{table}

\subsubsection{Raw Eigenvalues}
\label{sec:raw-eigs}

The time and space complexities of computing the full \(N \times N\)
correlation matrix of \(\mathbf{M}\) (and subsequent eigenvalues) are too
large to be tractable. However, since transposition does not change the
eigenvalues, we can use the transpose to efficiently compute these eigenvalues
(see Appendix \ref{sec:transpose}). These raw eigenvalues (feature \code{eigs}
in Table \ref{tab:features}), most directly test the idea that eigenvalues
alone are useful functional connectivity features.

However, not all eigenvalues may have predictive utility. We also examine as
features the central eigenvalues (defined as the middle 10\%, 20\%, or 40\% of
the sorted eigenvalues, \code{eigsmiddle} in Table \ref{tab:features}), and
tail eigenvalues (defined as the first and last 5\%,  10\%, and 20\% of each
tail of the spectrum, \code{eigsminmax} in Table \ref{tab:features}).

\subsubsection{RMT Features}
\label{sec:rmt-features}

To compare the observed eigenvalue distribution to some RMT theoretical
predictions, the eigenvalues must be unfolded
\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}. The
details and motivation behind the unfolding procedure are well documented in
\cite{guhrRandommatrixTheoriesQuantum1998a}. Practically, the unfolding process
can be viewed as a smoothing and rescaling procedure where the originally
observed spectrum \(\lambda_1 \le \dots \le \lambda_n\) is smoothly mapped to
an unfolded spectrum \(e_1, \dots, e_n\), such that \(\sum d_i/n \approx 1\),
if \(d_i = e_{i+1} - e_i \). For most empirical data, the true smoothing
function is not known, and so must be approximated
\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}. Typically,
a polynomial is chosen \cite{abul-magdUnfoldingSpectrumChaotic2014}.

Given empirically observed eigenvalues \(\Lambda\), the spectral rigidity,
\(\Delta_3(L)\), is calculated for any positive real value \(L <
\max(\Lambda)\) as:
\begin{equation}
\label{eq:rigidity}
\Delta_3(L) = \left \langle \min_{A,B} \frac{1}{L} \int_c^{c+L} \left(  \eta(\lambda) -A \lambda - B \right)^2 \right \rangle_c
\end{equation}
where \(\eta(\lambda)\) is the number of unfolded eigenvalues less than or
equal to \(\lambda\), \(\langle \cdot \rangle_c\) denotes the average with
respect to all starting points \(c\), and where \(A\) and \(B\) denote the
slope and intercept, respectively, of the least squares fit of a straight line
to \(\eta(\lambda)\) on \([c, c+L]\)
\cite{guhrRandommatrixTheoriesQuantum1998a}. Viewing the unfolded eigenvalues
as a timeseries, the spectral rigidity for a value \(L\) is the
\textit{average nonlinearity} of all intervals of length \(L\) over the series.

The level number variance, \(\Sigma^2(L)\), or \textit{level variance}, for short, is
closely related to the spectral rigidity \cite{mehtaRandomMatrices2004}, and is
calculated as:
\begin{equation}
\label{eq:levlvar}
\Sigma^2(L) = \left\langle \eta^2(L, c) \right\rangle_c - \left\langle \eta(L, c) \right\rangle^2_c
\end{equation}
where \(\eta(L, c)\) is the number of unfolded eigenvalues in \([c, c+ L]\),
and where \(c\), \(L\), and \(\langle \cdot \rangle_c\) are as above
\cite{guhrRandommatrixTheoriesQuantum1998a}. Viewing the unfolded eigenvalues
as an irregular timeseries, the level number variance is the
variation of the number of samples in all intervals of length \(L\) over the
series.

As eigenfeatures, we compute the unfolded eigenvalues, and the spectral
rigidity and level number variance for all \(L \in \{ 1, 2, \dots 20 \}\).
These are \code{unfolded}, \code{rigidity}, and \code{levelvar}, respectively,
in Table \ref{tab:features}.

\textit{Trimming Variants}. As the unfolding procedure operates on the sorted
eigenvalues, and involves fitting a smooth polynomial or exponential to these
values, extreme values are often omitted from fitting
\cite{moralesImprovedUnfoldingDetrending2011,
abueleninEffectUnfoldingSpectral2012}. The number of eigenvalues to trim is
subjective, and unfortunately, with a large number of different source
matrices, it is too destructive to naively use the same hard criterion (e.g.
largest 10\%, largest three) for all unfoldings.

We develop and test four trimming variants: no trimming, precision-based,
largest, and middle, with less subjective criteria for determining trimming
thresholds. The details of these trimming procedures are straightforward, and
can be found in Appendix \ref{sec:trimming}, or the
\href{https://github.com/DM-Berger/random-matrix-fmri/blob/7c9e4187f582dedee728cd7193b8894d928c2f00/code/rmt/updated_dataset.py#L431-L444}{source
code}.

\textit{Unfolding Variants}. Following any trimming, the eigenvalues are fit
with a polynomial. Because the choice of degree has been known to dramatically impact
certain analyses involving the spectral rigidity or level
variance\cite{abul-magdUnfoldingSpectrumChaotic2014,
moralesImprovedUnfoldingDetrending2011, abueleninSpectralUnfoldingChaotic2018,
fossionRandommatrixSpectraTime2013, abueleninEffectUnfoldingSpectral2012}, we
compute the rigidity and level variance with all possible combinations of
trimming and unfolding polynomial degrees of 3, 5, 7, and 9.

\textit{Software Release}. Because the computation of these statistics is non-trivial, and to make the
unfolding procedure more transparent and reproducible, we developed and release
the separate, open-source Python library
\code{empyricalRMT}\cite{dm-bergerStfxecutablesEmpyricalRMTV12022}. The library
allows for efficient, parallel computation of the spectral rigidity and level
variance via Monte Carlo methods, and automatically ensures convergence of the
statistics to user-specified tolerances. The library also makes available
various other functions useful for empirical RMT analyses, such as unfolding
and trimming functions, and plotting facilities for classic RMT ensembles.

\subsubsection{Smoothed Eigenfeatures}
\label{sec:eigs-only}

When computing RMT eigenfeatures, the polynomial fitting during unfolding has a
smoothing effect, as do the additional averaging \(\langle\;\rangle_c\)
operations. However, it is possible other, simpler transformations of the
eigenvalues (e.g. uniform or Savitsky-Golay smoothing) might have equal or
greater predictive utility. We thus also test smoothed variants of the
eigenvalues as predictive features: the sorted eigenvalues are smoothed with
either a uniform (moving average) filter or Savitsky-Golay, using window sizes
of 3, 5, 7, and 9 to yield 8 total features (\code{eigs\_smooth} and
\code{eigs\_savgol} in Table \ref{tab:features}).

\subsubsection{Feature Combinations}

A combination of RMT and non-RMT eigenfeatures could be more predictive than
either alone. We test this with a variety of eigenfeature combinations, with a
focus on combinations that involve the simplest features (e.g. raw eigenvalues,
unfolded eigenvalues) and then features involving additional processing (e.g.
smoothed eigenvalues, level variance, rigidity). Combined features are formed
by concatenation, so that if we have features \(f_1, \dots,  f_n\) with
dimensions \(p_1, \dots, p_n\), then the combined feature is \([f_1; \dots;
f_n]\) with \(\sum p_i\) dimensions.  The final combinations chosen can be
found in Table \ref{tab:features}, where the "\code{+}" symbol indicates concatenation.

\subsubsection{Slicing Variants}

The largest, middle, or smallest eigenvalues could be most useful in
characterizing some system. E.g. if the smallest eigenvalues correspond to
random / noise aspects of a system, but differences in the nature of the system
noise most differentiate between systems, then the smallest eigenvalues may
have the most predictive utility.  Likewise, the \(L\) value in each RMT
eigenfeature defines the degree of locality at which we summarize the spectrum:
with small \(L\)-values, the spectral rigidity summarizes the non-linearity of
eigenvalues that are relatively close to each other in magnitude. At large
values of \(L\), the rigidity summarizes the long-range non-linearity of the
spectrum.

Since predictive utility may vary with summary locality, or with the region of
the original spectrum, we investigate various front, middle, and end contiguous
slices of each eigenfeature in all analyses, where the size of each slice is
either the first or last 5\%, 10\%, or 20\% of the full eigenfeatures for
non-middle slices, or the middle 10\%, 20\%, or 40\% of the full eigenfeature for
the middle eigenvalues.

For combined features, slicing variants are also computed, with slicing
performed first on each feature to be combined. E.g. if we have features \(f_1,
\dots,  f_n\) with dimensions \(p_1, \dots, p_n\), then the sliced max 25\%
feature for each component feature, \(\hat{f}_i\) is the last \(p_i / 4\)
elements of \(f_i\), and the combined sliced feature is  \([\hat{f}_1;\;
\dots;\; \hat{f}_n]\)\footnote{Exact implementation can be found in the
\href{https://github.com/DM-Berger/random-matrix-fmri/blob/7c9e4187f582dedee728cd7193b8894d928c2f00/code/rmt/updated\_predict.py\#L182-L213}{source code}}.

\subsubsection{Baseline Features}
\label{sec:baselines}

If RMT- or eigenvalue-based features fail to predictively outperform features
that are simple summary statistics of the fMRI data, it is difficult to justify
the greater computational and interpretational complexities of the former. We
compute ten simple statistical reductions of the voxel time series ( "tseries"
features in Table \ref{tab:features}) to help assess the relative value of RMT
features.

The baseline reductions used as statistical summaries were: robust measures of
location (mean, max, min); non-robust measures of location (median,
\(95^{\text{th}}\) percentile, \(5^{\text{th}}\) percentile); non-robust
measures of scale (standard deviation and the range—\code{T-rng} in Table
\ref{tab:features}); and robust measures of scale (interquartile range and
difference between \(95^{\text{th}}\) percentile and \(5^{\text{th}}\)
percentile—\code{T-rrng} in Table \ref{tab:features}). Each statistic reduces
the fMRI data along the voxel dimension, yielding a feature of \(T\) dimensions
(e.g. the ``mean'' feature, \code{T-mean}, is the usual global mean signal).
Exact details can be found in the
\href{https://github.com/DM-Berger/random-matrix-fmri/blob/master/code/rmt/enumerables.py#L51-L69}{paper
source code}.

We do not take slice variants of these baseline features, since the baseline
features are still time series. However, we do also evaluate a number of
smoothing degrees of these features, to account for noise and to be somewhat
similar to the various smoothing variants for the eigenfeatures. Each baseline
feature is tested with a degree of uniform smoothing, where the size of the
smoothing window is either 1 (no smoothing), 2, 4, 8, or 16.

\subsection{Classifiers}

We use a variety of standard machine-learning classifiers available in the
Scikit-learn\cite{pedregosaScikitlearnMachineLearning2011} Python library to
solve each classification task. We use a gradient-boosted decision tree (GBDT),
random forest (RF) classifier, support vector classifier (SVC) with radial
basis function, and k-nearest neighbors classifiers with k equal to 3, 5, and 9
(KNN3, KNN5, and KNN9, respectively), in all cases with the default
hyperparameter values. We originally also attempted to test a simple logistic
regression classifier, but found that this model frequently failed to converge
for a number of features, even after increasing iterations significantly.

\subsection{Preprocessing Levels}

Each preprocessing intermediate in the (1) brain extraction, (2) slice time
correction, (3) motion correction, and (4) template registration pipeline was
saved and used for entirely separate feature analyses. E.g. all features were
extracted from an fMRI image prepared with one of four levels or degrees of
preprocessing, where preprocessing level \(k\) includes preprocessing steps 1 to \(k\),
inclusive, and starting at \(k = 1\).

\subsection{Normalization}

Because of the exponential distribution of the eigenfeatures, normalization is
somewhat non-trivial, and a simple standardization or min-max normalization is
ineffective. We instead first apply a logarithm to all eigenfeatures, and then
test each classifier with un-normalized or min-max normalized versions of the
log-features. Baseline features are also tested with un-normalized and min-max
normalized versions, but without any log transform.

\subsection{Multiverse Analysis}
\label{sec:multiverse}

We perform a \textit{multiverse
analysis}\cite{steegenIncreasingTransparencyMultiverse2016} to assess the
overall predictive potential of the various eigenfeatures across all
previously-mentioned analytic choices. That is, for each combination of
comparison task, classifier, and analytic choices, we use 5-fold cross
validation, and use the mean AUROC (mAUROC) across folds as our metric to
evaluate feature predictive utility.

The AUROC metric was chosen because it handles class imbalances present in the
various datasets, and is naturally normalized and interpretable such that
mAUROC \(<\) 0.5 indicates predictive performance worse than guessing, and
mAUROC \(>\) 0.5 indicates positive predictive
utility\cite{mandrekarReceiverOperatingCharacteristic2010}. However, we also
collected as performance metrics the accuracy, adjusted accuracy (accuracy
minus the proportion of the largest class), and the f1 score, and these metrics
are available with all the raw data for this study \ref{sec:data-and-code}.

the different

Each non-baseline feature is evaluated, for each classification task and
classifier, exactly 1280 times: there are 4 preprocessing levels \(\times\) 2
normalization methods \(\times\) 4 trimming choices \(\times\) 4 unfolding
degrees  \(\times\) 10 slice choices. Each baseline feature is evaluated
exactly 40 times for each comparison task and classifier: 4 preprocessing
levels \(\times\) 2 normalization methods \(\times\) 5 smoothing window sizes.
In total, with all the datasets examined in this study, this yields 2 053 920
mAUROC values to summarize, with 5 main \textit{analytic factors}—
preprocessing, normalization, trimming, unfolding/smoothing degree, and
slicing—to consider.

\section{Results}
\label{sec:results}

Summarizing this quantity of data requires some caution. Measures of location or
scale, even when robust, are, for the most part, misleading and uninformative
when distributions are skewed. In most cases we find skewed mAUROC
distributions, and so have chosen primarily to present our findings visually,
with kernel density estimates. However, we also provide overall numerical
summaries of the feature performances in Table \ref{tab:numerical}.

\begin{table}[h!]
\caption{Numerical summaries of feature mAUROCs across predictable comparisons,
and all combinations of analytic choices, sorted by 95\% percentile (robust
max) value. Bold values indicate column "best" values, when reasonable. We
note, however, that aggregating across datasets in this manner can be deeply
misleading\cite{brabecModelEvaluationNonconstant2020}, and thus the values here
are intended solely as supplements to distribution plots. }
\label{tab:numerical}
\small
\centering

\begin{tabular}{lrrrrrrr}
\hline
Feature &      mean &       min &        5\% &       50\% &       95\% &       max &       std \\
\hline
eigs + eigs\_smooth            &  \textbf{0.583} &  0.188 &  0.408 &  0.567 &  \textbf{0.788} &  0.906 &  0.111 \\
eigs + savgol                  &  0.580 &  0.188 &  0.408 &  \textbf{0.563} &  0.785 &  0.913 &  0.110 \\
unfolded                       &  0.568 &  0.150 &  0.400 &  0.555 &  0.775 &  0.931 &  0.108 \\
unfolded + levelvar            &  0.569 &  0.150 &  0.400 &  0.555 &  0.775 &  0.925 &  0.108 \\
unfolded + rigidity            &  0.568 &  0.150 &  0.401 &  0.555 &  0.775 &  0.925 &  0.108 \\
unfolded + rigidity + levelvar &  0.568 &  0.150 &  0.401 &  0.555 &  0.775 &  0.931 &  0.108 \\
eigs + rigidity + levelvar     &  0.564 &  0.202 &  0.401 &  0.548 &  0.763 &  \textbf{0.937} &  0.105 \\
eigs + unfolded                &  0.564 &  0.202 &  0.401 &  0.549 &  0.763 &  0.919 &  0.105 \\
eigs + unfolded + levelvar     &  0.564 &  0.202 &  0.400 &  0.548 &  0.763 &  0.933 &  0.105 \\
eigs + unfolded + rigidity     &  0.564 &  0.202 &  0.401 &  0.548 &  0.763 &  0.920 &  0.105 \\
eigs + levelvar                &  0.563 &  0.202 &  0.400 &  0.548 &  0.758 &  0.914 &  0.104 \\
eigs                           &  0.563 &  0.202 &  0.401 &  0.548 &  0.756 &  0.875 &  0.104 \\
eigs + rigidity                &  0.563 &  0.202 &  0.400 &  0.548 &  0.756 &  0.882 &  0.104 \\
eigs\_smooth                   &  0.569 &  0.188 &  0.417 &  0.556 &  0.755 &  0.907 &  0.102 \\
T-p05                          &  0.526 &  0.249 &  0.340 &  0.500 &  0.754 &  0.856 &  0.110 \\
eigs\_savgol                   &  0.566 &  0.188 &  0.410 &  0.553 &  0.748 &  0.909 &  0.102 \\
eigsminmax20                   &  0.556 &  0.245 &  0.411 &  0.544 &  0.738 &  0.906 &  0.098 \\
eigsminmax10                   &  0.550 &  0.231 &  0.409 &  0.538 &  0.737 &  0.906 &  0.095 \\
eigsminmax5                    &  0.549 &  0.213 &  0.400 &  0.539 &  0.729 &  0.888 &  0.096 \\
T-mean                         &  0.557 &  0.208 &  \textbf{0.427} &  0.558 &  0.717 &  0.814 &  0.085 \\
levelvar                       &  0.534 &  0.075 &  0.367 &  0.527 &  0.712 &  0.879 &  0.101 \\
eigsmiddle40                   &  0.547 &  \textbf{0.324} &  0.419 &  0.533 &  0.697 &  0.785 &  0.084 \\
rigidity + levelvar            &  0.529 &  0.154 &  0.377 &  0.523 &  0.696 &  0.888 &  0.094 \\
rigidity                       &  0.529 &  0.154 &  0.378 &  0.523 &  0.694 &  0.888 &  0.094 \\
T-rrng                         &  0.545 &  0.281 &  0.384 &  0.542 &  0.688 &  0.759 &  0.090 \\
eigsmiddle20                   &  0.545 &  0.321 &  0.419 &  0.533 &  0.686 &  0.798 &  0.080 \\
T-iqr                          &  0.494 &  0.175 &  0.312 &  0.490 &  0.682 &  0.759 &  0.099 \\
eigsmiddle10                   &  0.541 &  0.323 &  0.411 &  0.530 &  0.681 &  0.755 &  0.081 \\
T-std                          &  0.513 &  0.252 &  0.392 &  0.500 &  0.679 &  0.729 &  0.086 \\
T-rng                          &  0.538 &  0.279 &  0.392 &  0.539 &  0.678 &  0.779 &  0.092 \\
T-max                          &  0.540 &  0.275 &  0.385 &  0.546 &  0.678 &  0.754 &  0.095 \\
T-med                          &  0.539 &  0.221 &  0.399 &  0.541 &  0.665 &  0.702 &  0.077 \\
T-p95                          &  0.524 &  0.294 &  0.402 &  0.514 &  0.653 &  0.693 &  0.077 \\
T-min                          &  0.504 &  0.321 &  0.426 &  0.500 &  0.582 &  0.698 &  0.046 \\
\hline
\end{tabular}
\end{table}

In addition, each analytic factor could produce an interaction. For example, it
could be that trimming impacts the overall mAUROC distribution only at a
certain preprocessing level and for certain feature slicing. However, there are
too many potential interactions to present, so we limit our presentation to
main effects.  As there are a large number of features, we also summarize
patterns of results across two abstract groupings of similar features
("coarse", and "fine") shown in Table \ref{tab:features}.

It is also important to exclude a classification task or an analytic factor
from summaries if there is no evidence that the classification task is solvable
in general, or for that factor. We take as lack of evidence of solvability an
mAUROC distribution that is either (1) roughly symmetric and has mean and
median close to 0.5, or (2) with median and mode less than 0.5.

\begin{figure}
\begin{center}
\begin{tabular}{c}
\includegraphics[width=6.5in]{coarse_feature_overall_by_subgroup.png}
\end{tabular}
\end{center}
\caption
{ \label{fig:main-results}
AUROC distributions across gross feature groupings and comparison tasks.}
\end{figure}

\subsection{Classification Tasks}

\subsubsection{Non-Predictable Comparisons}
Based on the mAUROC distributions for either coarsely-grouped (Figure
\ref{fig:main-results}) or finely-grouped features (Figure
S\ref{S-fig:feature-group-all}), neither bilinguality nor depression were
predictable by either baseline or eigenfeatures. Additionally, the \code{pain v
duloxetine} comparison in the \textit{Osteo} dataset, and the
\code{trait\_attend v trait\_nonattend} conditions ("WeeklyAttend" in Figures
\ref{fig:main-results} and S\ref{S-fig:feature-group-all}) in the \textit{Attention}
dataset were also not meaningfully predictable by any feature. As such, we exclude
these classification tasks from further figures and discussion.

\subsubsection{Predictable Comparisons}
As visible in Figure \ref{fig:main-results}, when coarsely summarizing
features, eigenfeatures were more likely to be predictively useful than not,
and except for in the \code{task\_attend v task\_nonattend} comparison, were
also more predictively useful than the baseline features.  Eigenfeatures most
strongly and consistently demonstrated predictive utility in the \textit{Aging}
dataset \code{older v younger} classification task, the \textit{Osteo} dataset
\code{nopain v duloxetine} task, and in the \textit{Attention} \code{vigilant v
nonvigilant} comparison.

\begin{figure}
\begin{center}
\begin{tabular}{c}
\includegraphics[width=6.5in]{coarse_feature_by_preproc_predictive_subgroup.png}
\end{tabular}
\end{center}
\caption
{ \label{fig:preproc} mAUROC distributions by preprocessing degree.}
\end{figure}

\subsection{Largest mAUROCs}
Considering the various analytic choices as tunable parameters, it makes sense
to examine the \textit{largest} portion of mAUROCs as an indication of the
maximum predictive potential of the eigenfeatures. In this case, it is clear
that eigenfeatures using RMT features almost always had the highest potential
predictive utility (Figure \ref{fig:largest}). Figure S\ref{S-fig:fine-largest}
shows that this was primarily due to either the "rmt only" or "rmt + eigs"
features (see Table \ref{tab:features}). However, these features were also most
likely to cause poor performance and overfitting (indicated by an mAUROC \(<\)
0.5; Figure S\ref{S-fig:fine-smallest}).

Examining these RMT features more closely (Figure S\ref{S-fig:fine-degree}),
it is clear these features are more predictive mostly due to the unfolding
procedure: the mAUROC distributions with largest medians and modes are due to
features that include the unfolded eigenvalues. Combined features that use
the unfolded eigenvalues plus some other RMT eigenfeature have identical
mAUROC distributions to those using the unfolded eigenvalues alone.

\begin{figure}
\begin{center}
\begin{tabular}{c}
\includegraphics[width=6.5in]{coarse_feature_largest_by_subgroup.png}
\end{tabular}
\end{center}
\caption
{ \label{fig:largest} Distributions of largest 500 mAUROCs by coarse feature grouping.}
\end{figure}

\subsection{Effect of Preprocessing}
At a coarse level, slice time correction followed by motion correction tended
to slightly increase the predictive utility of the eigenfeatures (Figure
\ref{fig:preproc}) relative to brain extraction only. Subsequent registration
following these steps did not generally impact mAUROC distributions further,
except in the \code{task\_attend v task\_nonattend} comparison, where
registration reduced the predictive utility of the eigenfeatures (Figure
\ref{fig:preproc}, second-last column). When examining features more finely, it
is clear that preprocessing most impacts the mAUROC distribution of the largest
and central eigenvalues ("eigs middle" and "eigs max" in Figure
S\ref{S-fig:fine-preproc}).

\subsection{Effect of Classifier}
mAUROC distributions were generally similar across coarse or fine feature
groupings (Figure S\ref{S-fig:coarse-classifier}, S\ref{S-fig:fine-classifier}).
In the \textit{Attention} \code{task\_attend v task\_nonattend} condition, eigenfeatures were modally predictive only with
the random forest classifier, whereas in the \textit{Aging} data, a SVC was least
likely to have mAUROC \(<\) 0.5. Overall, however, differences in the mAUROC distributions
due to classifier were small.

\subsection{Normalization}
There was no visible effect of feature normalization on mAUROC distributions.

\subsection{Effect of Trimming}
Trimming based on numerical precision (see Section \ref{sec:rmt-features} and
Appendix \ref{sec:trimming}), did not result in meaningfully different mAUROC
distributions in any case (Figure S\ref{S-fig:fine-trim}). However, trimming
away the largest, or both largest and smallest eigenvalues, generally had a
significant positive effect on the predictive quality of RMT features, most
especially for RMT features involving the unfolded eigenvalues. When employing
these trimming methods, these features were consistently more predictive than
not (Figure S\ref{S-fig:fine-trim}).

\subsection{Effect of Unfolding Degree}
The choice of polynomial unfolding degree significantly impacted the mAUROC
distributions for most classification tasks and most RMT features, and most
significantly for the level variance features (Figure
S\ref{S-fig:fine-degree}). Overall, Figure S\ref{S-fig:fine-degree} weakly
suggests that either smaller (degree = 3) or larger (degree = 9) unfolding
degrees tend to yield the most predictively useful RMT features. However, when
restricting to the most predictive RMT features (those including the unfolded
eigenvalues), it seems clear from Figure S\ref{S-fig:fine-degree} the the
largest unfolding degree of 9 produces the most favourable mAUROC
distributions.

\subsection{Effect of Slicing}
Features were slightly more predictive when using either the largest 5 or 10\% or middle
25\% ("max-05", "max-10" and "mid-25" in Figure
\ref{S-fig:best-params-slicing}), and least predictive when features primarily
involved the smallest eigenvalues.

\section{Discussion}

In this study, eigenfeatures inspired by RMT and derived from the eigenvalues
of the full, whole-brain voxelwise fMRI correlation matrix, were found to have
predictive utility across a wide variety of phenomena and analytic choices.
Compared to simple baseline reductions of the fMRI data, these eigenfeatures
had more consistent predictive utility, and a higher maximum predictive
potential (Figures \ref{fig:main-results}-\ref{fig:largest}). In addition to
evidence from previous studies
\cite{sebaRandomMatrixAnalysis2003,wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020}
this suggests RMT may be a useful analytic and theoretical tool for
understanding functional connectivity.

\begin{figure}
\begin{center}
\begin{tabular}{c}
\includegraphics[width=6.5in]{unfolded_osteo_duloxetine_v_nopain.png}
\end{tabular}
\end{center}
\caption
{ \label{fig:unfolded} Raw eigenvalues and unfolded eigenvalues (polynomial
degree 9) for \textit{Osteo} dataset \footnotesize\texttt{duloxetine v nopain}
classification task, brain extraction and slice time correction only. Median
observed mAUROC = 0.643 (range = 0.346 - 0.925)}
\end{figure}

It is likely that the unfolded eigenvalues have predictive utility because of
their smoothing and re-scaling effect. Figure \ref{fig:unfolded}, which depicts
the raw eigenvalues and unfolded eigenvalues with different trimming
procedures, shows how the raw eigenvalues have strongly exponential
distributions, even with logarithmic axes. This is due to the
magnitude of the largest eigenvalues, and the trimmed feature distributions in
Figure \ref{fig:unfolded} are less pathologically distributed.

Also, as is clear from Table \ref{tab:numerical} and Figure S\ref{S-fig:fine-classifier}, that
a simple combination of the raw eigenvalues plus smoothed versions of them are the most
generally-predictive features examined.

It was also sometimes the case that adding RMT-specific features (e.g. the spectral
rigidity, or level variance) to either the unfolded eigenvalues, or raw
eigenvalues, would allow for a higher maximum mAUROC (Table
\ref{tab:numerical}, Figure S\ref{S-fig:fine-largest}). One case where these
features yielded the largest predictions was in the \textit{Aging} dataset.
This is depicted in Figure \ref{fig:observables}, where it is clear that, after
trimming, at almost any \(L\) value, subgroups are strongly and clearly
visually-separable.

\begin{figure}
\begin{center}
\begin{tabular}{c}
\includegraphics[width=6.5in]{observables_older_younger_v_older.png}
\end{tabular}
\end{center}
\caption
{ \label{fig:observables} Rigidity and level variance features (unfolding
polynomial degree 3) for \textit{Aging} dataset \footnotesize\texttt{younger v
older} classification task, brain extraction only. Median observed mAUROC =
0.703 (range = 0.424 - 0.880). Horizontal axis L-values range from 1 to 20 in
all subplots, as described in the \ref{sec:methods}.}
\end{figure}

However, the range of mAUROCs observed in all cases in this study suggests
caution. All eigenfeatures and baselines features were sensitive to analytic
choices such as the amount of preprocessing, feature slicing (including trimming) and
feature smoothing / unfolding, and this sensitivity was such that all features
resulted in mAUROCs on either side of 0.5. Most of these analytic choices are benign enough that
it is difficult to rigorously justify one choice over another, meaning RMT-based
analyses may be susceptible to post-hoc justification \cite{kerrHARKingHypothesizingResults1998}.

It is also difficult to quantify the practical significance of the observed
mAUROC values. With small datasets of rs- or task-fMRI data, binary
classification using whole-brain features is challenging. For subject level
classification, state of the art deep learning methods improve upon guessing by
17\% for autism \cite{bengs4DSpatioTemporalDeep2020} or 3-30\% for ADHD
\cite{riazDeepFMRIEndtoendDeep2020}, 16\% for severe depression
\cite{ramasubbuAccuracyAutomatedClassification2016}, and 23\% for obsessive
compulsive disorder \cite{takagiNeuralMarkerObsessiveCompulsive2017}. Manual
feature engineering with more separable conditions (e.g. schizophrenia) can
result in classification accuracies well above 90\%
\cite{duHighClassificationAccuracy2012}, and with larger data, sophisticated
custom feature extraction methods can achieve near perfect accuracies at
classifying task vs. rest
\cite{zhangCharacterizingDifferentiatingTaskbased2016}. However, for functional
connectivity data and with classical machine learning algorithms (such as SVM)
we in general only expect large prediction accuracies (e.g. greater than 80\%)
when the group functional connectivities are already strongly separated (e.g.
Cohen’s \(d > 1.0\); \citenum{dansereauStatisticalPowerPrediction2017}). In
this study eigenfeatures could improve upon guessing by 16.2\% - 33.7\%
(\ref{S-tab:acc-numerical}).

It is somewhat surprising that eigenfeatures are, on net, predictively-useful
at all. The reduction of the functional connectivity matrix to \(T - 1\)
eigenvalues is aggressive, and the subsequent small-degree polynomial fit used
in the unfolding procedure further reduces information in the raw data.
Still, the subsequent features more often than not allow non-trivial 5-fold
cross-validated mean prediction AUROCs or accuracies.

We believe that future studies using RMT to inform their examinations be more
careful. In the absence of a cross-validated predictive framework that takes
into account the strong impact of researcher analytic choices, any apparent
explanatory utility of RMT should be treated with skepticism. RMT deals with
highly-ideal \textit{iid} matrices, but almost no real-world systems involve
\textit{iid} elements. The basic RMT insight that eigenvalues can be used to
describe matrix-representable systems is certainly vindicated here, but the use
of mathematically non-trivial RMT statistics (like the spectral rigidity or
level number variance) should not be used if simpler statistical descriptions
of the eigenvalues are equally informative.

\section{Conclusion}

\section{Disclosures}

\section{Acknowledgments and Funding}

\section{Data, Materials, and Code Availability}
\label{sec:data-and-code}

All raw fMRI data used in this study is publicly available on the
\href{https://openneuro.org/}{OpenNeuro}
platform\cite{poldrackOpenfMRIOpenSharing2017,
markiewiczOpenNeuroResourceSharing2021}.

All code used specifically to perform preprocessing and analyses for this paper
is publicly available \href{https://github.com/DM-Berger/random-matrix-fmri/}{online} in
the paper GitHub repository\cite{dm-bergerDMBergerRandommatrixfmriPaper2022}.
This repository also includes all produced prediction data in a single table
available in the file \code{all\_produced\_prediction\_data.json.xz} available
\href{https://github.com/DM-Berger/random-matrix-fmri/blob/master/all\_produced\_prediction\_data.json.xz}{in
the paper repository}.

Code used to compute the spectral observables and
perform unfolding is publicly available in the
\href{https://github.com/stfxecutables/empyricalRMT}\code{empyricalRMT}
library\cite{dm-bergerStfxecutablesEmpyricalRMTV12022}.

% \subsection{Page Setup and Fonts}

% All text and figures, including footnotes, must fit inside a text area 6.5 in.\
% wide by 9 in.\ high (16.51 by 22.86 cm). Manuscripts must be formatted for US
% letter paper, on which the margins should be 1 in.\ (2.54 cm) on the top, 1
% in.\ on the bottom, and 1 in.\ on the left and right.

% The Times New Roman font is used throughout the manuscript, in the sizes and
% styles shown in Table~\ref{tab:fonts}. If this font is not available, use a
% similar serif font. The manuscript should not contain headers or footers. Pages
% should be numbered.

% \begin{table}[ht]
% \caption{Fonts sizes and styles.}
% \label{tab:fonts}
% \begin{center}
% \begin{tabular}{|l|l|} %% this creates two columns
% %% |l|l| to left justify each column entry
% %% |c|c| to center each column entry
% %% use of \rule[]{}{} below opens up each row
% \hline
% \rule[-1ex]{0pt}{3.5ex}  Document entity & Brief description  \\
% \hline\hline
% \rule[-1ex]{0pt}{3.5ex}  Article title & 16 pt., bold, left justified  \\
% \hline
% \rule[-1ex]{0pt}{3.5ex}  Author names & 12 pt., bold, left justified   \\
% \hline
% \rule[-1ex]{0pt}{3.5ex}  Author affiliations & 10 pt., left justified   \\
% \hline
% \rule[-1ex]{0pt}{3.5ex}  Abstract & 10 pt.  \\
% \hline
% \rule[-1ex]{0pt}{3.5ex}  Keywords & 10 pt.  \\
% \hline
% \rule[-1ex]{0pt}{3.5ex}  Section heading & 12 pt., bold, left justified  \\
% \hline
% \rule[-1ex]{0pt}{3.5ex}  Subsection heading & 12 pt., italic, left justified  \\
% \hline
% \rule[-1ex]{0pt}{3.5ex}  Sub-subsection heading & 11 pt., italic, left justified  \\
% \hline
% \rule[-1ex]{0pt}{3.5ex}  Normal text & 12 pt. \\
% \hline
% \rule[-1ex]{0pt}{3.5ex}  Figure and table captions &  10 pt. \\
% \hline
% \end{tabular}
% \end{center}
% \end{table}

% \section{Parts of Manuscript}

% This section describes the normal structure of a manuscript and how each part
% should be handled. The appropriate vertical spacing between various parts of
% this document is achieved in LaTeX through the proper use of defined
% constructs, such as \verb|\section{}|.

% \subsection{Title and Author Information}
% \label{sect:title}
% The article title appears left justified at the top of the first page. The
% title font is 16 pt., bold. The rules for capitalizing the title are the same
% as for sentences; only the first word, proper nouns, and acronyms should be
% capitalized. Do not begin titles with articles (for example, a, an, the) or
% prepositions (for example, on, by, etc.). The word ``novel'' should not appear
% in the title, as publication will imply novelty. Avoid the use of acronyms in
% the title, unless they are widely understood. Appendix A contains more about
% acronyms.

% The list of authors immediately follows the title, 18 points below. The font is
% 12 pt., bold and the author names are left justified. The author affiliations
% and addresses follow the names, in 10-pt., normal font and left justified. For
% multiple affiliations, each affiliation should appear on a separate line.
% Superscript letters (a, b, c, etc.) should be used to associate multiple
% authors with their respective affiliations. The corresponding author should be
% identified with an asterisk, and that person's email address should be provided
% below the keywords.

% \subsection{Abstract}
% The abstract should be a summary of the paper and not an introduction. Because
% the abstract may be used in abstracting journals, it should be self-contained
% (i.e., no numerical references) and substantive in nature, presenting concisely
% the objectives, methodology used, results obtained, and their significance.
% Please note that the following journals require the use of structured abstracts
% in manuscript submissions: \textit{Neurophotonics}, the \textit{Journal of
% Biomedical Optics}, and the\textit{ Journal of Medical Imaging}. Structured
% abstracts are encouraged for the \textit{Journal of Micro/Nanolithography,
% MEMS, and MOEMS}. Helpful guidelines for structured abstracts are available on
% the website of the journal.

% \subsection{Subject terms/Keywords}
% Keywords are required. Please provide 3-6 keywords related to your paper.

% \subsection{Body of Paper}
% The body of the paper consists of numbered sections that present the main
% findings. These sections should be organized to best present the material.

% To provide transition elements in your paper, it is important to refer back (or
% forward) to specific sections. Such references are made by indicating the
% section number, for example, ``In Sec.\ 2 we showed...'' or ``Section 2.1
% contained a description...'' If the word Section, Reference, Equation, or
% Figure starts a sentence, it is spelled out. When occurring in the middle of a
% sentence, these words are abbreviated Sec., Ref., Eq., and Fig.

% At the first occurrence of an acronym, spell it out followed by the acronym in
% parentheses, for example, charge-coupled diode (CCD).

% \subsection{Footnotes}
% Textual footnotes should be used rarely to present important documentary or
% explanatory material whose inclusion in the text would be
% distracting.\footnote{Example of a footnote.} Due to problems with HTML
% display, use of footnotes should generally be avoided. If absolutely necessary,
% the footnote mark must come at the end of a sentence. To insert a footnote, use
% the {\verb|\footnote{}|} command.

% \subsection{Appendices}
% Brief appendices may be included when necessary, such as derivations of
% equations, proofs of theorems, and details of algorithms. Equations and figures
% appearing in appendices should continue sequential numbering from earlier in
% the paper.

\subsection{Disclosures}
Conflicts of interest should be declared under a separate header, above
Acknowledgments. If the authors have no competing interests to declare, then a
statement should be included declaring no conflicts of interest. For assistance
generating a disclosure statement, see the form available from  the
International Committee of Medical Journal Editors website:
\linkable{http://www.icmje.org/conflicts-of-interest/}

\subsection{Acknowledgments}
Acknowledgments and funding information should be added after the conclusion,
and before references. Include grant numbers and the full name of the funding
body. The acknowledgments section does not have a section number.

\subsection{Data, Materials, and Code Availability}
Availability of data, materials, and/or code used in the research results
reported in the manuscript may be declared under the heading ``Data, Materials,
and Code Availability,'' following the Acknowledgments section. As relevant,
provide specific access information or restrictions for data, materials, and
computer code (i.e., links to repository access addresses with guidance on
commercial or public access).

% \subsection{References}
% The References section lists books, articles, and reports that are cited in the
% paper. This section does not have a section number. The references are numbered
% in the order in which they are cited. Examples of the format to be followed are
% given at the end of this document.

% The reference list at the end of this document is created using BibTeX, which
% looks through the file {\ttfamily report.bib} for the entries cited in the
% LaTeX source file.  The format of the reference list is determined by the
% bibliography style file {\ttfamily spiejour.bst}, as specified in the \\
% \verb|\bibliographystyle{spiejour}| command.  Alternatively, the references may
% be directly formatted in the LaTeX source file.

% For books\citenum{Lamport94,Alred03,Goossens97} the listing includes the list
% of authors (initials plus last name), book title (in italics), page or chapter
% numbers, publisher, city, and year of publication.  Journal-article references
% \citenum{Metropolis53,Harris06} include the author list, title of the article
% (in quotes), journal name (in italics, properly abbreviated), volume number (in
% bold), inclusive page numbers or citation identifier, and year.  A reference to
% a proceedings paper or a chapter in an edited book\citenum{Gull89a} includes
% the author list, title of the article (in quotes), conference name (in
% italics), editors (if appropriate), volume title (in italics), volume number if
% applicable (in bold), inclusive page numbers, publisher, city, and year.
% References to an article in the SPIE Proceedings may include the conference
% name, as shown in Ref.~\citenum{Hanson93c}.

% The references are numbered in the order of their first citation. Citations to
% the references are made using superscripts, as demonstrated in the preceding
% paragraph. One may also directly refer to a reference within the text, for
% example, ``as shown in Ref.~\citenum{Metropolis53} ...''  Two or more
% references should be separated by a comma with no space between them. Multiple
% sequential references should be displayed with a dash between the first and
% last numbers \citenum{Alred03,Perelman97,Lamport94,Goossens97,Metropolis53}.

\subsubsection{Reference linking and DOIs}
A Digital Object Identifier (DOI) is a unique alphanumeric string assigned to a
digital object, such as a journal article or a book chapter, that provides a
persistent link to its location on the internet. The use of DOIs allows readers
to easily access cited articles. Authors should include the DOI at the end of
each reference in brackets if a DOI is available. See examples at the end of
this manuscript. A free DOI lookup service is available from CrossRef at
\\\linkable{http://www.crossref.org/freeTextQuery/}. The inclusion of DOIs will
facilitate reference linking and is highly recommended.

In the present LaTeX template, the author needs to add the DOI reference by
including it in a ``note'' in the bibliography file, as shown in the file
{\verb+report.bib+}, for example, \\ {\verb+note = "[doi:10.1117/12.154577]"+}.
The DOI may be used by the reader to locate that document with the link:
{\verb+http://dx.doi.org10.1117/12.154577+}.

\subsection{Biographies}
A brief professional biography of approximately 75 words may be provided for
each author, if available. Biographies should be placed at the end of the
paper, after the references. Personal information such as hobbies or
birthplace/birthdate should not be included. Author photographs are not
published.

% \section{Section Formatting}
% \label{sect:sections}
% In LaTeX, a new section is created with the \verb|\section{}| command, which
% automatically numbers the sections. Sections will be numbered sequentially,
% starting with the first section after the abstract, except for the
% acknowledgments and references. (Note that numbering of section headings is not
% required, but the numbering must be consistent if used.) All section headings
% should be left justified.

% Main section headings are in 12-pt. bold font, left-justified and in title
% case, where important words are capitalized.

% Paragraphs that immediately follow a section heading are leading paragraphs and
% should not be indented, according to standard publishing style. The same goes
% for leading paragraphs of subsections and sub-subsections. Subsequent
% paragraphs are standard paragraphs, with 0.2-in (5 mm) indentation. There is no
% additional space between paragraphs. In LaTeX, paragraphs are separated by
% blank lines in the source file. Indentation of the first line of a paragraph
% may be avoided by starting it with \verb|\noindent|.

% \subsection{Subsection Headings}
% All important words in a subsection (level 1) header are capitalized.
% Subsection numbers consist of the section number, followed by a period, and the
% subsection number within that section, without a period at the end. The heading
% is left justified and its font is 12 pt. italic.

% \subsubsection{Sub-subsection headings}
% The first word of a sub-subsection is capitalized. The rest of the text is not
% capitalized, except for proper names and acronyms (the latter should only be
% used if well known). The heading is left justified and its font is 11 pt.
% italic.

% \section{Figures and Tables}

% \subsection{Figures}

% Figures are numbered in the order in which they are called out in the text.
% They should appear in the document in numerical order and as close as possible
% to their first reference in the text. It may be necessary to move figures or
% tables around to enhance readability. LaTeX will attempt to place figures at
% the top or bottom of a page in which they are first referenced.

% Figures, along with their captions, should be separated from the main text by
% 0.2 in.\ or 5 mm and centered. Figure captions are centered below the figure or
% graph. Figure captions start with the abbreviation ``Fig'' in front of the
% figure number, followed by a period, and the text in 10-pt. font. See
% Fig.~\ref{fig:example} for an example.

% \begin{figure}
% \begin{center}
% \begin{tabular}{c}
% \includegraphics[height=5.5cm]{mcr3b.eps}
% \end{tabular}
% \end{center}
% \caption
% { \label{fig:example}
% Example of a figure caption. }
% \end{figure}

% Authors may wish to create figures consisting of two or more images, in which
% case, they should be neatly arranged in a rectangular array.  In no case,
% should the article's text be wrapped around a figure. Figure~\ref{fig:example2}
% shows two side-by-side images. When a figure contains more than one image, the
% author must submit them as a single image file. Further details about figure
% formatting can be found in the author guidelines for each specific SPIE
% journal: \\

% \linkable {https://www.spiedigitallibrary.org/journals/journal-authors}.

% \begin{figure}
% \begin{center}
% \begin{tabular}{c}
% \includegraphics[height=5.5cm]{fig2.eps}  % fig2 includes two images
% \\
% (a) \hspace{5.1cm} (b)
% \end{tabular}
% \end{center}
% \caption
% { \label{fig:example2} Example of a figure containing multiple images: (a) sun
% and (b) blob. Figures containing multiple images must be submitted to SPIE as a
% single image file.}
% \end{figure}

% \subsection{Tables}
% Tables are numbered in the order in which they are referenced. They should appear in the document in numerical order and as close as possible to their first reference in the text. It is preferable to have tables appear at the top or bottom of the page, if possible. Table captions are handled identically to those for figures, except that they appear above the table. See Table~\ref{tab:fonts} for an example.

% \subsection{Multimedia}
% Acceptable file formats, including MOV (.mov), MPEG (.mpg), and MP4 (.mp4), are playable using standard media players, such as VLC or Windows Media Player. The recommended maximum size for each multimedia file is 10-12 MB. Authors must insert a representative still image from the video file in the manuscript as a figure. The caption label will be linked by the publisher to the actual video file. The video may also be mentioned in an existing figure caption. Multimedia files are treated in the same manner as figures and they will be numbered sequentially with normal figures.  The video number, file type, and file size should be included in parentheses at the end of the figure caption. See Figure \ref{vid:satellite} for an example.

% \begin{video}
% \begin{center}
% {\includegraphics[height=5cm]{satellite.eps}}
% \\
% \end{center}
% \caption{\label{vid:satellite}This satellite is a still image from Video 1 (Video 1, MPEG, 2.5 MB).}
% \end{video}

\appendix    % this command starts appendixes

\section{Correlation Eigenvalues via Transposition}
\label{sec:transpose}

Let \(\mathbf{X}\) is a real \(n \times p\) matrix with \(n > p\). Let
\[
\mathbf{Z} = \text{norm}(\mathbf{X}) =
\left(
\mathbf{X}_1 - \bar{\mathbf{X}}_1 | \dots | \mathbf{X}_p - \bar{\mathbf{X}}_p
\right)
\]
where \(\mathbf{X}_i\) denotes column \(i\) of \(\mathbf{X}\). Denote the (unordered) set
of eigenvalues of \(\mathbf{X}\) as \(\text{eigs}(\mathbf{X})\), and let \(r = (p - 1)^{-1}\).
Denote the covariance matrix of \(\mathbf{X}\) as \(\text{cov}(\mathbf{X})\).  Then:

\begin{align*}
\text{eigs}\left( \text{cov}(\mathbf{X})  \right)
&= \text{eigs}(r \cdot \mathbf{Z}\mathbf{Z}^{\top}) \\
&= r \cdot \text{eigs}(\mathbf{Z}\mathbf{Z}^{\top}) \\
&= r \cdot \text{eigs}((\mathbf{Z}\mathbf{Z}^{\top})^{\top}) \\
&= r \cdot \text{eigs}(\mathbf{Z}^{\top}\mathbf{Z}) \\
\end{align*}

Denote the correlation matrix of \(\mathbf{X}\) as \(\text{corr}(\mathbf{X})\), and let
\[
\mathbf{Y} = \text{stand}(\mathbf{X}) =
\left(
(\mathbf{X}_1 - \bar{\mathbf{X}}_1) / \sigma_1 | \dots | (\mathbf{X}_p - \bar{\mathbf{X}}_p) / \sigma_p
\right)
\]

Then

\[
\text{eigs}\left( \text{corr}(\mathbf{X})  \right)
= \text{eigs}\left( \text{cov}(\mathbf{Y})  \right)
= r \cdot \text{eigs}(\mathbf{Y}^{\top}\mathbf{Y})
\]

\section{Trimming Procedures}
\label{sec:trimming}

We implement three trimming procedures: precision-based, largest, and middle
trimming. The
\href{https://github.com/DM-Berger/random-matrix-fmri/blob/7c9e4187f582dedee728cd7193b8894d928c2f00/code/rmt/updated_dataset.py#L431-L444}{source
code} is the definitive reference for the procedures, but we describe the
motivations belows.

In precision-based trimming, we trim away any eigenvalues that are close enough
to zero to be considered a result of numerical error due to floating point
representation. There are two thresholds we consider, including the one used by
NumPy\cite{harrisArrayProgrammingNumPy2020} for the
\href{https://numpy.org/doc/stable/reference/generated/numpy.linalg.matrix_rank.html}{determination
of matrix rank}, and those recommended by LAPACK\cite{laug} in their user
guide\cite{andersonLAPACKUsersGuide1999a} on the
\href{https://netlib.org/lapack/lug/node89.html}{error bounds for symmetric
eigenproblems}, and related
\href{https://netlib.org/lapack/lug/node90.html}{additional details}. We trim
each matrix eigenvalues to whichever threshold is largest for the matrix in
question.

For largest trimming, we must determine a threshold at which to separate
"large" from "small" eigenvalues. The eigenvalues for our data tended to grow
exponentially, so we instead looked at thresholding on the logarithms. We then
used \(k\)-means with \(k=2\) on the precision-trimmed eigenvalues, and took
the largest cluster (which also always had the smaller mean) cluster as the
"largest" eigenvalues to trim away. "Middle" trimming simply reflects the
threshold found by the largest trim method, e.g. if the largest trim method
removes the last \(n\) precision-trimmed eigenvalues, then we also trim the
first \(n\) smallest eigenvalues remaining after precision-trimming.

We chose one-dimensional k-means partly due to efficiency and simplicity, and
because of the general relation to classical thresholding methods like the Otsu
method\cite{liuOtsuMethodKmeans2009}.

% \section{Miscellaneous Formatting Details}
% \label{sect:misc}
% At times it may be desired, for formatting reasons, to break a line without starting a new paragraph. In a LaTeX source file, a linebreak is created with \verb|\\|.

% \subsection{Formatting Equations}
% Equations may appear inline with the text, if they are simple, short, and not of major importance; for example, $\beta = b/r$.  Important equations appear on their own line.  Such equations are centered.  For example, ``The expression for the field of view is
% \begin{equation}
% \label{eq:fov}
% 2 a = \frac{(b + 1)}{3c} \, ,
% \end{equation}
% where $a$ is the ...''  Principal equations are numbered, with the equation number placed within parentheses and right justified.

% Equations are considered to be part of a sentence and should be punctuated accordingly. In the above example, a comma appears after the equation because the next line is a subordinate clause. If the equation ends the sentence, a period should follow the equation. The line following an equation should not be indented unless it is meant to start a new paragraph. Indentation after an equation is avoided in LaTeX by not leaving a blank line between the equation and the subsequent text.

% References to equations include the equation number in parentheses, for example, ``Equation~(\ref{eq:fov}) shows ...'' or ``Combining Eqs.~(2) and (3), we obtain...'' Note that the word ``Equation'' is spelled out if it begins a sentence, but is abbreviated as ``Eq.'' otherwise. Using a tilde in the LaTeX source file between two characters avoids unwanted line breaks, for example between ``Eq.'' and the following equation number..

% \subsection{Formatting Theorems}

% To include theorems in a formal way, the theorem identification should appear in a 10-point, bold font, left justified, and followed by a period.  The text of the theorem continues on the same line in normal, 10-pt. font, achieved in LaTeX using \verb|\footnotesize|.  For example,

% \vspace{2ex}\noindent{\footnotesize\textbf{Theorem 1.} For any unbiased estimator...}

% \disclosures
\subsection*{Disclosures}
Conflicts of interest should be declared under a separate header. If the authors have no relevant financial interests in the manuscript and no other potential conflicts of interest to disclose, a statement to this effect should also be included in the manuscript.

\subsection* {Acknowledgments}
This unnumbered section is used to identify those who have aided the authors in understanding or accomplishing the work presented and to acknowledge sources of funding.

\subsection* {Data, Materials, and Code Availability}
As relevant, the availability of data, materials, and/or software code used in the research results reported in the manuscript may be declared in this section. (Note: this section is required for the \textit{Journal of Biomedical Optics} and \textit{Neurophotonics}.) Provide specific access information or restrictions for data, materials, and computer code (i.e., links to repository access addresses with guidance on commercial or public access).

%%%%% References %%%%%

\bibliography{report}   % bibliography data in report.bib
\bibliographystyle{spiejour}   % makes bibtex use spiejour.bst

%%%%% Biographies of authors %%%%%

\vspace{2ex}\noindent\textbf{First Author} is an assistant professor at the University of Optical Engineering. He received his BS and MS degrees in physics from the University of Optics in 1985 and 1987, respectively, and his PhD degree in optics from the Institute of Technology in 1991.  He is the author of more than 50 journal papers and has written three book chapters. His current research interests include optical interconnects, holography, and optoelectronic systems. He is a member of SPIE.

\vspace{1ex}
\noindent Biographies and photographs of the other authors are not available.

\listoffigures
\listoftables

\end{spacing}
\end{document}